### 2023-06-16 body.tex
Initial Overleaf Import

\maketitle

\tableofcontents
\listoftheorems[ignoreall,show={definition,corollary,lemma,theorem}]

\section{Introduction}

Deleting content from Swarm is meant to impose a cost on retrieval of deleted content that is no less than the cost of having the entirety of the information stored from deleting to attempted retrieval. Clearly, the imposed cost cannot be greater than  this, because by storing the information (either in Swarm or on one's own drive or someplace else) always allows one to access it. Note that for those who did not access the information before it was deleted this cost is usually prohibitive.

\subsection{Access control vs. censorship}

Deleting content or revoking access to content are difficult notions. On the one hand, the digitisation and increasing ubiquity of personal data in the cloud touch on sensitivities more than ever before. On the other hand, cheap storage makes it virtually impossible to make material once seen become unseen.
In response to the former, regulators often formulate requirements on removability that fails to recognise the latter. The regulatory appeal of centralised gate-keeping is clear: in case of violation, the gate-keeper provides a  well-defined responsible party. Be it on the level of publishing platform or hosting infrastructure, the  operating entity is in control of the technology serving the content and thus lends itself as the target of legal enforcement when it comes to denying access. Such a clear course of action gives the unwary a phony sense of security when it comes to digital publishing while the real issue is swept under the rug: publishing platforms have the technological means to filter content. However, such centralised control always provides an attack surface, since it makes exerting influence on content cheaper.  What starts out as sensible measures of content curation with time becomes extant censorship. This is all the more problematic with social platforms that gained quasi-monopolistic status due to network effects. As a result of high switching cost, innocent content creators sometimes end up being 'deplatformed'. Similarly, the ability to identify hosts and deny access to content via legal decree gives powers to be the means to infringe on freedom of speech.

In the decentralised paradigm of web 3, there is no longer an operating entity behind either the publishing platform or the hosting infrastructure. Therefore, the cost of censorship is raised beyond the realm of feasibility. However, a novel approach is imminent to replace the sense of security gate-keepers purportedly represent. We argue that technology that allows sovereign control over one's own data is needed.

\subsection{Deletion and revoking access}

First, notice that any reference to removal of information in the sense of erasing it from all physical storage devices is both unenforceable and impractical. Even the strictest data protection audits do not require the erasure of offending data from backup tapes. In general, the tacit assumption is that information ordered to be removed become inaccessible through \emph{typical precedented methods of access}.

In this paper, we aspire to formulate the strongest meaningful definition of deletion applicable to decentralised storage and offer a construction that implements it.
Importantly, such a construction needs to be purely technical, referring to primary actors' capabilities and costs as opposed to procedural, referring to intermediaries' obligations to respect the rights of primary actors.

Primary actors here are the \emph{uploader} that wishes to share content by granting read access to a number of parties, called \emph{downloaders}. Deletion is simply reformulated as revoking access
to the content from all parties including oneself. Granting access is defined as providing some canonical \emph{reference} to the content that the system can eventually resolve to the full information meant to be disclosed. Any party that is privileged to access this information, is able to store, re-code and potentially disseminate the content seen at any later point in time including whatever process would qualify as deleting or removing access. There is, by definition, no protection possible against such adversity, hence retrospective revocation of access (and therefore deletion) is meant to be used in a narrower sense. Practically, the notion of revoked access needs to be defined in terms of \emph{the viability to replay the same access method} or any action that would allow the adversary to access information after their access is revoked for a cost lower than the cost of storing information totaling the entire size of the content shared for the entire period starting at the time the content is shared.

The scheme for content with access revocation has the following requirements:

\begin{itemize}
    \item \emph{specialisation} - Uploader is able to choose at the time of publishing a specialised construct allowing retrospective revocation of access.
    \item \emph{sovereignity} -- Uploader is the sovereign owner of the data and is in sole possession of the means to selectively revoke access from any party previously granted access.
    \item \emph{security} -- After revocation of access, grantee is unable to access the content with the reference granting access, or any other cue that incurs less cost than the cost of storing the information in its entirety.
\end{itemize}

From a user's perspective, content that is meant to be reliably deleted should be uploaded as such. The costs of uploading such content is allowed to be a (small integer) multiple of the cost of regular, censorship-resistant uploads.

The uploader's credentials are necessary to delete their own deletable content. Deletables must also allow access control, i.e., only available to a specific set of recipients.

The point of deleting is that if the downloader has NOT stored at least as much information as the deleted information itself, they will have no way of retrieving it. In particular, retaining the reference used to download the content in the first place will no longer be sufficient after it has been deleted. However a stronger statement is that no other information shorter than the deleted content will suffice.

\subsection{Immutable storage and access through references}

Swarm's core layer is a peer to peer overlay network that implements a distributed storage model called \emph{DISC (distributed immutable store of chunks)}, see figure \ref{fig:disc}).  Some information is said to be available in Swarm if one can retrieve it using a \emph{reference} that is (much) shorter than the information itself and is in fact constant in the length.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/disc.pdf}
  \caption{DISC}
\label{fig:disc}
\end{figure}

For example, a \emph{Swarm hash}, or a Swarm hash with a \emph{decryption key}  or a \lstinline{bzz:} link.  These references are collision-free content addresses and therefore are immutable.%
%
\footnote{Even is we use mutable references, e.g., an ENS address, or feed (based on single owner chunks), these can be resolved at the time to a content address.  Whoever obtains this information can store the resolved primary reference which can later be accessed as long as assumptions about data retention in the blockchain and/or Swarm hold.} 
%
Swarm is a decentralized network where individual participants cannot (and should not) be trusted to play by the rules; it is the built-in incentives that encourage cooperative behavior and help maintain it, it being a Nash-equilibrium. The condition for content deletion to be achieved is that in \emph{at least one neighborhood} all nodes are honest, i.e., do not operate a malicious node which would store and serve information that they should not store and serve. With the growth of the Swarm network, the number of neighborhoods increases whereas the membership of neighborhoods remains about the same. Thus, the costs of attacking the proposed scheme are expected to go up.

One cooperative neighborhood suffices, that is. Of course, a powerful adversary can infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it!) and keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand. But there is a huge price on such indiscriminate archiving of all Swarm's content and that is the only way reliably defeat the cryptographic construction that we propose for deleting.

Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and it is not accessed sufficiently often, the chunks constituting it will be garbage collected. However, these cannot be reliably deleted in the above defined sense.

\section{Construction}

The goal of this section is to arrive at a formal construction of a DISC-based revocable access model. We will restrict our scheme to \emph{chunks}, the fundamental fix sized storage unit of Swarm's DISC model.  The proposed \emph{dream} construct implements a  deletable content storage and access model conforming to the requirements of specialisation, sovereignty, and security. The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model. On top of this, as customary in swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} of access control.

\begin{itemize}
    \item[\textbf{d}] \emph{deniable} -- the dream key serves as a one time pad for decryption. Since multiple content chunks (in fact any arbitrary content) can use the same dream pad, the key's association to any content is plausibly deniable.
    \item[\textbf{r}] \emph{revocable} -- access granted by dream keys is revocable. Revoking access from all parties including oneself is taken as deletion.   
    \item[\textbf{e}] \emph{expirable} -- the scheme allows for one-time use, i.e., the key can only be retrieved once
    \item[\textbf{a}] \emph{addressable} -- access can be granted to a neighbourhood, only clients operating a node with an overlay address in a given range  are able to access
    \item[\textbf{m}] \emph{malleable}  --  the construct is resilient to churn and dynamic change of network size, is reusable across independent grantees and upgradeable.

\end{itemize}

The scheme is built on top of DISC's APIs and can be implemented entirely as a second layer solution.  Despite the rich feature set, the scheme is not a complex cryptographic scheme, rather relies on the interplay of various levels and component subsystems.

\begin{itemize}
    \item \emph{one time pad} -- Dream content is reconstructed by xor-ing two equal-length parts. This scheme is shared by a construct designed  for plausibly deniable content.
    \item \emph{authorised mutability} -- finding a scheme where retrieval will rely on an interactive protocol.
    \item \emph{incentives} -- the smooth operation of the interactive protocol is guaranteed by the systemic incentives, however, correctness should not depend on them. Abuse of the intention, should be prohibitively costly to incentivise. 
\end{itemize}

\subsection{Prerequisites}

\subsubsection{Chunk upload in swarm}

When content is uploaded to swarm, the local client splits the data into 4-kilobyte chunks and pushes each chunk to the network using the \emph{pushsync} protocol. This protocol is responsible for transmitting a chunk newly entered to the neighbourhood where it will be retrieved, i.e., the neighbourhood closest to the content address of the chunk. See figure \ref{fig:pushsync}.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/push-sync.pdf}
  \caption{Push syncing.}
\label{fig:pushsync}
\end{figure}

Network protocol in swarm use forwarding Kademlia routing meaning that queries reach their target neighbourhood by being relayed through a series of hops by nodes with ever increasing proximity to the target address.

The chunk delivery to storer nodes can also be seen as a request: the closest node respond with a statement of custody that is passed back via the same route as the original request (\emph{backwarding}). 
When uploading content, a measure must be present that would prevent mass injection of random chunks to the network. By requiring \emph{postage stamps}, we impose an upfront cost to uploads which is an effective measure against spamming, i.e., makes it costly to flood the network with chunks and as a consequence expunge valuable chunks from the network by triggering garbage collection on nodes with  capacity shortage. The attached postage stamp can serve as a signal to the  importance of keeping a chunk.

As an extention to this idea, the central postage stamp contract operates a system of redistribution called \emph{postage lottery}  whereby the pool of funds cumulating from all postage revenue is reallocated to unstaked storer nodes in return for proof of custody on chunks with valid stamps.

This system is responsible for positive storage incentives. More relavant to push syncing, the promise of storage revenue from the postage lottery race bestows a value on a chunk with a postage stamp and renders the pushsynced chunks a valueable commodity for the nodes close to them.
An intermediate node is incentivised to forward a chunk towards its address because the node it passes the chunk on to will pay more for it than what the node buys it for. As an important corollary to this, a node will always have the incentive to push a chunk if they do not need to pay for it or if they can make sure they are compensated.

\begin{lemma}[uploaded chunks travel to neighbourhood]
\label{lem:pushsync}
Every chunk uploaded to swarm is pushsynced and is forwarded by each relaying node towards its address  provided it has a valid postage stamp attached. As a consequence of Kademlia connectivity, the route to the node whose address is closest to the chunk can be found using local decisions by the nodes intermediate in the path. As a result of incentivised push syncing, a chunk with a valid postage stamp will end up being received by at least the closest node. Due to synchronisation of chunks among the nodes in a neighbourhood, the chunk will end up being received by each node in the neighbourhood of its address.
\end{lemma}

\subsubsection{Single owner chunks}

The DISC storage model supports two chunk types, a content addressed chunk and  single owner chunk. Content addressed chunks attain their integrity by the fact that their address is deterministically derived from their content with the help of a one-way uniform digest function (in particular, the binary  merkle tree hash). A single owner chunk on the other hand, encodes the association of an index and a payload sanctioned by the signature of an owner. The owner's ethereum address hashed with the index determines the chunk address. See figure \ref{fig:soc}.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/single-owner-chunk.pdf}
  \caption{Single owner chunks}
\label{fig:soc}
\end{figure}

Single owner chunks can wrap full length  content addressed chunks.

The hash function used to generate the content address of chunks in swarm is based on a \emph{binary merkle tree}. For ease of notation though, we just use $\mathit{H}$ to indicate a hash function. This is also appropriate since the particular choice of hash function does not affect the correctness   of our construction.

\subsubsection{Pseudo-random sequence from generator}
Assume we have a pseudo-random deterministic function that generates a longer (e.g., chunk-sized = 4K) sequence $c$ from a key-sized (32 byte) generator $g$. For example, the Keccak sponge function used throughout Ethereum for hashing does have this capability. It is used in Swarm in order not to introduce  security assumptions in addition to those of Ethereum.

\begin{definition}[Pseudo-random sequence generation]
\label{def:gen4k}
\begin{equation}
\mathit{Gen4K}: [32]byte \mapsto [4096]byte
\end{equation}

\begin{equation}
\mathit{Gen4K}(g) \defeq \langle c_0, c_1, \ldots c_{127}\rangle
\end{equation}
with

\begin{equation}
c_i \defeq \mathit{H}(g|i)
\end{equation}
\end{definition}

\subsection{Dream: deniable, revocable, expirable, addressable, malleable}

In what follows, through a number of steps, we introduce the \emph{dream}, a chunk-sized one time pad construction that provides the necessary mutability for content with revocable access.

Let $B$ be a set  of postage batches, owned by $U$. Without losing generality, we assume that this set is represented by a postage payment ID $b$, a 32-byte hash.

Let \emph{Chunks(b)} denote all chunks stamped by $b$.

\begin{definition}[Nearest Chunk]
\label{def:cin}
Define  $\mathit{NC}(b,p)$ as the chunk stamped by batch $b$ with minimal XOR distance to pivot $p$.

\begin{equation}
\mathit{NC}: \mathit{Batches}\times \mathit{Chunks} \mapsto \mathit{Chunks}
\end{equation}
\begin{equation}
\mathit{NC}(b,p)\defeq \underset{c\in \mathit{Chunks}(b)}{\operatorname{arg\,min}}\chi(\mathit{H}(p),\mathit{H}(c))
\end{equation}
\end{definition}

Let us define the \emph{chunk filter} as a function that transforms an input one time pad into another output one time pad using an additional chunk.

\begin{definition}[Chunk filter function]
\label{def:chunk-filter}
\begin{equation}
f: \mathit{chunks}\times\mathit{chunks}\mapsto \mathit{chunks}
\end{equation}
such that $f$(p,C) is calculated as follows:  
\begin{equation}
    f(p,C) \defeq \mathit{Gen4K}(\mathit{H}(p|C)) 
\end{equation}
...
\end{definition}

\begin{corollary}[Chunk filter is a random oracle]
\label{cor:chunk-filter-oracle}
Follows directly from the random oracle assumption about the hash function.
\end{corollary}

Let us define the \emph{task function $T$} for the dream construct. The task function defines an one time pad transform function using the chunk filter function on a set of chunks stamped with $b$ and found in the neighbourhood of the input $p$.

\begin{definition}[Task function]
\label{def:task}
Given a batch $b$, the task function takes the nearest chunk stamped by $b$ in the  neighbourhood of the input $p$.
\begin{equation}
T: \mathit{Batches} \mapsto (\mathit{Chunks} \mapsto \mathit{Chunks})
\end{equation}
such that
\begin{equation}
\mathit{T}(b)(p)\defeq f(p,\mathit{NC}(b,p)) 
\end{equation}

\end{definition}

\begin{corollary}[Task function is a random oracle]
\label{cor:task-function-oracle}
Due to \ref{cor:chunk-filter-oracle} and the choice of input $p$, the task function serves as an effective pseudo random generator function.
\end{corollary}

\begin{corollary}[Task function local computability]
\label{cor:task-function-locality}
The task function can be calculated by storer nodes in the neighbourhood of the input. This is a consequence of the DISC storage model, specifically that nodes store chunks in the proximity of their overlay node address.

In particular, if a storer node with overlay address $a$ stores all chunks $c_i$ where $\mathit{PO}(a,\mathit{H}(c_i))\geq d$ for some depth $d$ and $\mathit{PO}(a,\mathit{H}(p))\geq d$ as well for pivot $p$ then if it stores at least one chunk from batch $b$, it also stores $NC(b,p)$, because it implies that $a$, $\mathit{H}(p)$ and $\mathit{H}(NC(b,p))$ have a common prefix of length $d$.

If, however, $\mathit{PO}(a,\mathit{H}(NC(b,p)))<d$ then the computation of the task function fails, as the node does not store any chunks from $b$.

As long as all storer nodes store at least one chunk from $b$ and the entire address space is covered with storer nodes, the task function is computable by the storer node with overlay address $\underset{a\in S}{\operatorname{arg\,min}}\chi(\mathit{H}(p), a)$ where $S$ is the set of all storer nodes.
In particular the $\mathit{PO}(p,\mathit{NC}(b,p))$ should always be greater than the storer nodes neighbourhood depth. Conversely, if swarm grows and the density of storer nodes is so high that there is a node whose neighbourhood  depth is higher than the the minimum PO used for nearest chunks, then computability of the task function cannot be guaranteed.

This practically means that the dream imposes a minimum density requirement on batches. Let us use the term network depth for this minumum value.

%This reduces to proving that such a node has all the chunks in $\mathit{CiN}(b,d,h)$. 
%If the value  of chunk in  $b$ is   higher than the cut off value, then the node should not have deleted any chunk in $\mathit{chunks}(b)$.% 
% 
%\footnote{This translates to the requirement that batch  $b$ has sufficient chunk balance as witnessed on the blockchain.}   
%
%As long as it  had received all chunks it will have all chunks within its storage radius $r$, $\mathit{CiN}(b,r,o)$. We need to show that the chunks needed  to calculate the task function are a subset of this, i.e., 
%$\mathit{CiN}(b,d,h) \subseteq \mathit{CiN}(b,r,o)$.

%Since $o$  is closest to $h$, $PO(o,h)\geq r$.
%Therefore $\mathit{CiN}(b,r,o) =  \mathit{CiN}(b,r,h)$.
%As $d>r$ by  construction, $\mathit{CiN}(b,d,h)\subseteq %\mathit{CiN}(b,r,h)$.  \qedsymbol
\end{corollary}

Since the task function can be applied to its own output, we can define the one time pad chain function using a generic \emph{chaining} construct:

\begin{definition}[Chaining]
\label{def:chaining}
\begin{equation}
\mathit{Chain}: (X \mapsto X) \times X \mapsto (\mathbb{N} \mapsto X)
\end{equation}
such that
\begin{equation}
\mathit{Chain}(f,g)(i)\defeq\begin{cases}g&\text{ if }  i=0\\
f(\mathit{Chain}(f,g)(i-1)) & \text{otherwise}
\end{cases}
\end{equation}
\end{definition}

This allows us to define the \emph{dream chain}.

\begin{definition}[Dream chain]
\begin{equation}
\mathit{DreamChain}: \ma  thit{Batches}\times \mathit{Keys} \mapsto (\mathbb{N} \mapsto \mathit{Chunks})
\end{equation}
such that
\begin{equation}
\mathit{DreamChain}(b,g) \defeq \mathit{Chain}(T(b),\mathit{Gen4k}(g))
\end{equation}
\end{definition}

\begin{corollary}[Dream chain as pseudo-random generator]
\label{cor:dream-chain-oracle}
Due to \ref{cor:task-function-oracle}, the dream chain function is a pseudo random generator.
\end{corollary}
Now we can turn to the definition of a dream.

\begin{definition}[Deniable, revocable, expirable, addressable, malleable access method]
\label{def:dream-otp}
For step-count $n$ and authorised target address $a$ and a batch $b$ with sufficient density for network depth $d$, find the generator $g$, such that
the $n$-th one time pad output of the dream chain hashes to $a$'s neighbourhood using the same network depth $d$ that constrains nearest chunk.

\begin{equation}
\mathit{dream}(b,n,a) \subset \mathit{Keys}\times\mathit{Chunks}
\end{equation}
such that
\begin{equation}
\langle g, k\rangle \in \mathit{dream}(b,n,a) \Longleftrightarrow
k=\mathit{DreamChain}(b,g)(n)
\land
\mathit{PO}(a,H_{\mathit{bmt}}(k))\geq d
\end{equation}

\end{definition}

\begin{lemma}[Construction by batch owner]

The owner of the batch is able to construct the dream pad.

\begin{proof}
First, it is trivial to see that the owner of  $b$ has access to $\mathit{chunks(b)}$ and  therefore can  calculate $T(b)$ for any input. As a result they can also calculate $k$ given $g$. Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is $2^{-d}$, and thus the expected number of cycles to find the dream is $2^d$.\qed
\end{proof}
\end{lemma}

Let us  define a network protocol \emph{dream}. See figure \ref{fig:protocol}.

\begin{figure}[htbp]
  \centering
    % \includegraphics[width=0.8\textwidth]{}
  \caption{The dream protocol}
\label{fig:protocol}
\end{figure}

\begin{definition}[Dream protocol]

Those who play the protocol, listen to single owner chunks such that $c\in\mathit{dreamable}$.
We assume that they extract the batch identifier $b$, and the payload $p$ as input to the task function and then calculate $T(b)(p)$, wrap it as a dream chunk and send it to the network.
\end{definition}

\begin{corollary}[Reception of one time pad]
\label{cor:reception}
The one time pad in the chain is received by the node closest to its address.

\begin{proof}
Follows from lemma \ref{lem:pushsync} and the wrapping construction. \qedsymbol
\end{proof}
\end{corollary}

\begin{corollary}[Continuity of the chain]
\label{cor:continuity}
Given a chain  $c=\mathit{DreamChain}(b,g)$, 
a storer node at address $o$ closest to $\mathit{H}(c(i))$ is able to calculate $c(i+1)$.

\begin{proof}
Follows from \ref{cor:task-function-locality}.
\qedsymbol
\end{proof}
\end{corollary}

In other words, the neighborhood from which to choose the chunks to calculate $T$ is within the node's area of responsibility,  and therefore the node can safely assume that it does have all the  chunks.

\begin{corollary}[Termination]
\label{cor:termination}
Given $\langle g, k\rangle \in \mathit{dream}(b,g,a)$, the chain  $c=\mathit{DreamChain}(b,g)$, 
if the protocol is followed up to $n$ steps, then $k=c(n)$ is received by the target node at address $a$.

\begin{proof}
Follows from \ref{cor:reception} and \ref{cor:continuity}.\qedsymbol
\end{proof}
\end{corollary}

\subsection{Contructing and referencing dream chunks}

We now turn to the construction of dream chunks.

\begin{definition}[Constructing a dream chunk]
\label{def:dream}
Uploader $U$ wishes to grant downloader $D$ (at overlay address $a$) revocable  access to content chunk $C$.
$U$  chooses a postage batch $b$ it owns that is completely filled, a step-count $n$, and a depth $d$.
$U$ creates a dream generator $\langle g, k\rangle\in\mathit{dream}(b,d,n,a)$.

Now $U$ calculates the "ciphertext" $C'=C \xor k$ and uploads it to swarm obtaining 
$r=\mathit{H}(C')$ as reference.

Now $U$ creates the \emph{dream reference} as $\mathit{ref}(C)=\langle r,b,g\rangle $.

%This will be serialised as concatenated 97 bytes.
% $U$ gives the reference to $D$ who operates 
% a node at overlay address $a$.
\end{definition}

\begin{definition}[Retrieval]
\label{def:retrieval}
$D$ in possession of a dream chunk reference as 
$\langle r,b,g\rangle $ calculated by $U$  as $\mathit{ref}(C)$ constructs $p=\mathit{Gen4K}(g)$ and wraps it as a dream chunk and sends it to swarm. When it receives the dream chunk for $b$, it extracts payload $k$.

Then $D$ downloads $C'$ by using reference $r$, then decodes $C=C'\xor k$.
\end{definition}

\begin{lemma}[Retrieval is correct and viable]
The retrieval process described in \ref{def:retrieval} is trivially correct as long as 
\begin{enumerate}
    \item  $C'$ is retrievable through $r$. This 
    \item the dream protocol is adhered to
    \item $\mathit{NC}(b,p)$ has not changed for the neighbourhoods that are on the dream path. ($\mathit{chunks(b)}$) can only change if the owner chooses to add new chunks to the  batch which is exactly the tool to revoke access.
\end{enumerate}
\end{lemma}
is able to access $C$.

\begin{corollary}[Deniable dreams]
\label{cor:deniable}
The construction is deniable.

\begin{proof}
Beside our sensitive content $C$, take $A$, any uncontroversial content chunk. When producing $C'$, the owner also produces  $A'=A\xor k$ and upload it. When asked about $k$, producing $A$ makes the denial of other content including $C$ more plausible. \qedsymbol
\end{proof}
\end{corollary}

\begin{theorem}[Access control is correct]
\label{the:accessm}
As long as $\mathit{chunks}(b)$ is unchanged, $D$ launches a retrieval as per \ref{def:retrieval}, they will indeed receive chunk $k$ at address $a$. Once they can access content $C$.

Notice that  $D$ is unable to access $C$ at any other neighbourhood.
\begin{proof}
Follows from \ref{cor:termination}.   \qedsymbol
\end{proof}
\end{theorem}

We are now turning to access revocation. Deletion of content is equivalent to revoking all access.

\begin{definition}[revoking access to a dream chunk]
\label{def:revoking}

Uploader $U$ had granted $D$ (at address $a$) access to $C$ through the dream reference $\langle r,b,g\rangle $. $U$ revokes access from $D$ by uploading extra chunks to batch $b$ such that $\mathit{NC}$ changes.
\end{definition}

\begin{theorem}[revoking access is correct]
\label{the:revoking}

As long as there is a single honest neighbourhood walked by the protocol, such that the nearest chunk to pivot $p$ from batch $b$ changes, then $D$ can no longer retrieve $C$.

\begin{proof}
Let us assume that there is an $i$ such that $\mathit{NC}(b,H(c(i)))$ changed since uploader constructed the dream. Now, when the node in the neighbourhood of $H(c(i))$ calculates the task function, the result will be completely different (see \ref{cor:dream-chain-oracle}). \qedsymbol
\end{proof}
\end{theorem}

\section{Analysis}

\subsection{Incentives}

Discuss the reliability of the protocol in the context of economic incentives.
\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/push-payment.pdf}
  \caption{Incentives for push syncing.}
\label{fig:push-payment}
\end{figure}

How to determine when to forward.

discuss the cost of a dream
\subsection{Security}

We now turn to the discussion how to calibrate the step-count given a security model. Let us assume a network-wide neighbourhood infiltration rate of $\frac{1}{2}$, meaning that  half of the neighbourhoods in the network is assumed  to be  malicious and colluding.

Assuming that when  revoking access, the owner uploaded new chunks in each neighbourhood on the dream path. We assume that malicious nodes are able and willing to disregard these  new chunks and in an attempt to facilitate abuse of revoked access respond to the dream protocol just like  before the revocation.
If each  neighbourhood  on  the path comes from $M$, it is in principle possible that the dream protocol gives the unchanged response, and access is abused, i.e., allowed even though it is revoked.
However even if even  a single neighbourhood on the dream path is honest they respect the newly arrived chunks and divert the dream path, making it never terminate with the grantee.

Thus for  a breach to happen, all neighbourhoods should come from $M$.
Which neighbourhood is malicious is not known, therefore the choice of neighbourhood in the construction is considered independent. when choosing $n$ nodes, the probability of each being malicious is $2^{-n}$. So for an infiltration of $\frac{1}{2}$, and a security requirement of $99.9\%$, $n=10$, for $99.9999\%$, $n=20$ will suffice. For an infiltration of $\frac{1}{32}$, and a security requirement of $99.9\%$, $n=2$, for $99.9999\%$, $n=4$ will suffice.

Note that the reference does not leak either the step-count or any other information about the neighbourhoods touched. Moreover, participants in the protocol also do not know either at which position they are on the chain or anything  about the other neighbourhoods except for the following one to which they are push-syncing their chunk.

Discuss ways nodes could collude to circumvent the intention of deletion

\subsection{Dream files}

Generalise from chunks to files.

\section{Conclusion}

% Since these parts are symmetrical, it is somewhat misleading to refer to them as ciphertext and key.  However, in the deletable construct one part will be available via a normal hash reference, this we can call one time pad, while the ephemeral part is called encoded content.

% There is also a variation that "self-destructs", i.e. that deletes itself upon
% retrieval, but it is important to understand that it is a very fragile
% mechanism with no guarantee that the intended recipient will be able to
% download it even once. It should only be used in settings where the recipient can,
% for a limited time, request re-uploading in case the content gets deleted
% without all its parts reaching the recipient.

% \begin{definition}[pss encode function]

% \begin{equation}
% \Psi(b,d,c)=\mathit{Trojan}(\mathit{nonce}, \mathit{topic}, \mathit{payload})
% \end{equation}
% where
% \begin{equation}
% \mathit{nonce} = c[0:32]
% \end{equation}
% \begin{equation}
% \mathit{topic} = \tau
% \end{equation}
% \begin{equation}
% \mathit{payload}[0:32] =  b
% \end{equation}
% \begin{equation}
% \mathit{payload}[32:34] = \mathit{bigendian.Uint16}(d)
% \end{equation}
% \begin{equation}
% \mathit{payload}[34:4030] = c[100:4096]
% \end{equation}
% \end{definition}

### 2023-06-16 dream.tex
Initial Overleaf Import

\section{Dream}

The increasing trend of digitisation and the ensuing ubiquity of personal data in the cloud touch on sensitivities more than ever before. 
In response to the former, regulators often formulate requirements on removability that fails to recognise 
that it is virtually impossible to make material once seen become unseen.
Entities that operate infrastructure underpinning digital publishing are in control of the technology serving the content and  are able to implement "removal" by denying access.
Such centralised gate-keeping  has its regulatory appeal as it can be thought of as an instrument of law enforcement. Although such intervention rarely solves the breach of unrealistic privacy rights, the gate-keeper provides a well-defined responsible party with the effective course of action: not serving censored content.
While this gives the unwary a phony sense of security, it only exemplifies the bigger problem of censorship: publishing platforms have the technological means to filter content and centralised control makes exerting influence on content cheaper.  What starts out as sensible measures of content curation with time becomes extant censorship. This is all the more problematic with social platforms that gained quasi-monopolistic status due to netwtork effects. As a result of high switching cost, innocent content creators ever more often end up being 'deplatformed'. Similarly, the ability to identify hosts and deny access to content via legal decree gives powers to be the means to infringe on freedom of speech.

In the decentralised paradigm of web 3, there is no longer an operating entity behind either the publishing platform or the hosting infrastructure. Therefore, the cost of censorship is raised beyond the realm of feasibility. And yet, the potential permanent exposure of personal data is perceived as scary to most and some novel solution is called for to restore the sense of security gate-keepers purportedly represent.

\subsection{Deletion and revoking access}

First, notice that any reference to removal of information in the sense of erasing it from all physical storage devices is both unenforceable and impractical. Even the strictest data protection audits do not require the erasure of offending data from backup tapes. In general, the tacit assumption is that information ordered to be removed should become inaccessible through \emph{typical precedented methods of access}.

In what follows we formulate the strongest meaningful definition of deletion applicable to decentralised storage and offer a construction that implements it.
Importantly, such a construction is purely technical, referring to primary actors' capabilities and costs as opposed to procedural, referring to intermediaries' obligations to respect the rights of primary actors.

Primary actors here are the \emph{uploader} that wishes to share content by granting read access to a number of parties, called \emph{downloaders}.  Granting access is defined as providing some canonical \emph{reference} to the content that the system can eventually resolve to the full information meant to be disclosed. Any party that is privileged to access this information, is able to store, re-code and potentially disseminate it. This provides a myriad of ways to make accessible the content at any later point in time to get round whatever process would qualify as deletion or removing access. 
There is, by definition, no protection possible against such adversity. As a consequence, any legally sound notion of deletion (retrospective revocation of access) is meant to be interpreted in a narrower sense: \emph{the viability to replay the same access method}  for a cost lower than the \emph{full cost} of storing the content.%
%
\footnote{That is, the total cost of storage paid for the full size for the entire period kstarting from the time access was revoked uptill the attempted breach.}

Let us now define deletion as a scheme for uploading content with access revocation that has the following requirements:

\begin{itemize}[noitemsep]
    \item[--] \emph{specialisation} -- Uploader is able to choose at the time of publishing a specialised construct%
%
\footnote{From a user's perspective, content that is meant to be reliably deleted should be uploaded as such. The costs of uploading such content is allowed to be a (small integer) multiple of the cost of regular, censorship-resistant but non-deletable uploads.}
%
    allowing retrospective revocation of access.
    \item[--] \emph{sovereignty} -- Uploader is the sovereign owner of the data and is in sole possession of the means to selectively revoke access from any party previously granted access.%
%
\footnote{The uploader's credentials are necessary to delete their own deletable content. Deletables must also allow access control, i.e., only available to a specific set of recipients.}
%
    \item[--] \emph{security} -- After revocation of access, a grantee is unable to access the content using the same reference that the access was granted with, or any other cue shorter than the deleted content.%
%
\footnote{In other words, if the downloader has \emph{not} stored at least as much information as the deleted information itself, they will have no way of retrieving it.}
%
\end{itemize}

Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and it is not accessed sufficiently often, the chunks constituting it will be garbage collected. However, chunks with  expired postage stamps cannot be regarded as reliably deleted to satisfy the requirements of sovereignty and security.

\subsection{Construction}

The goal of this section is to arrive at a formal construction of a DISC-based revocable access model. We will restrict our scheme to \emph{chunks}, the fundamental fix sized storage unit of Swarm's DISC model.  The proposed \emph{dream} construct implements a  deletable content storage and access model conforming to the requirements of specialisation, sovereignty, and security. The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model. On top of this, as customary in swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} expressing requirements of access control.

\begin{itemize}
    \item[\textbf{D}] \emph{deniable} -- the dream key serves as a one time pad for decryption. Since multiple content chunks (in fact any arbitrary content) can use the same dream pad, the key's association to any content is plausibly deniable.
    \item[\textbf{R}] \emph{revocable} -- access granted by dream keys is revocable. Revoking access from all parties including oneself is taken as deletion.   
    \item[\textbf{E}] \emph{expirable} -- the scheme allows for one-time use, i.e., the key can only be retrieved once.
    \item[\textbf{A}] \emph{addressable} -- access can be granted to a neighbourhood, only clients operating a node with an overlay address in a given range  are able to access.
    \item[\textbf{M}] \emph{malleable}  --  the construct is resilient to churn and dynamic change of network size, is reusable across independent grantees and upgradeable.
\end{itemize}

The scheme is built on top of DISC's APIs and can be implemented entirely as a second layer solution.  Despite its rich feature set, the scheme does not use complex cryptography, but relies on the interplay of various component subsystems.

% The network protocols of Swarm use forwarding Kademlia  meaning that requests are routed to their destination through a series of hops successively relayed by forwarding nodes with ever increasing proximity to the target address.

% The chunk delivery to storer nodes can also be seen as a request: the closest node respond with a statement of custody that is passed back via the same route as the original request (\emph{backwarding}).

% \subsubsection{Pseudo-random sequence from generator}

Assume we have a pseudo-random deterministic function that generates a longer (e.g., chunk-sized = 4K) sequence $c$ from a key-sized (32 byte) generator $g$.%
%
\footnote{For example, the Keccak sponge function used throughout Ethereum for hashing does have this capability. Alternatively, the blockcipher encryption using initial nonce $g$ could be applied to a  fixed constant chunk such as all zeros.}

% \subsubsection{Dream: deniable, revocable, expirable, addressable, malleable}

The central construct called a \emph{dream} is a chunk-sized one time pad which acts as a decryption key for the deletable content. The dream chunk is collectively created by a set of nodes following a network protocol. Each node in the sequence of participating peers receives a piece of data and uses it as input together with data present in their reserve to produce an output sent to the neighbourhood of the next node in the chain. As long as each node in the chain performs the same calculation and forwards the same result, the same one time pad can be construed and ultimately the deletable content can be retrieved and read.

The insight here is that retrieving a deletable chunk's references involves calculations using a set of immutable chunks under the control of the uploader. The uploader's ability to change this underlying set holds the key to providing the necessary mutability required by any notion of deletion.

Let $b$ be the batch ID (a 32-byte hash) of a postage batch owned by $U$ and 
let $\mathit{Chunks}(\gamma,b,p)$ stand for all chunks stamped by $b$ at block $\gamma$ in the bucket designated by the pivot $p$. Define  $\overline{\Huge\chi}(\gamma,b,p)$ as the chunk stamped by batch $b$ in bucket designated by $p$ older than $\gamma$  whose address is closest (has minimal XOR distance) to pivot $p$.

\begin{eqnarray}
\overline{\Huge\chi}&:& \Gamma\times\mathit{Batches}\times \mathit{Segment} \mapsto \mathit{Chunks}\\
\overline{\Huge\chi}(\gamma,b,p)&\defeq &\underset{c\in \mathit{Chunks}(\gamma,b,p)}{\operatorname{arg\,min}}{\Large\chi}(p,\textsc{Address}(c))
\end{eqnarray}

If batch $b$ is underutilised, a bucket designated by $p$
may not contain any chunks belonging to the batch bucket.
In such cases there is no chunk closest to $p$, thus the \emph{BBB} function is undefined.

Let us now define the OTP update function $\Delta(\gamma,b)$
for batch ID and input address $p$ as follows: 
%
\begin{eqnarray}
{\Delta}&:&\Gamma\times\mathit{Batches}\mapsto\mathit{Chunks}\mapsto\mathit{Chunks}\\\
{\Delta}(\gamma, b)&:&\mathit{Chunks}\mapsto\mathit{Chunks}\\
\Delta(\gamma,b)(c)&\defeq&\mathcal{G}\idx{4K}(\textsc{Address}(\overline{\Huge\chi}(\gamma,b,c)))
\end{eqnarray}

Since uniformity depth of batches is meant to be larger than nodes' storage depth, all chunks in a batch bucket designated with $p$ is  in the reserve of every node in the neighbourhood designated by $p$.

If swarm grows and  there are nodes whose storage  depth is higher than the batch's uniformity depth, then the neighbourhood designated by $p$ may not contain all the
chunks in the bucket, thus compromising the computability of the OTP update function.

As long as 1) the bucket of batch $b$ designated by $p$ is not empty and 2) the neighbourhoods include (one or more) buckets then the update function is computable by any storer node within the neighbourhood designated by $p$.

Since the task function can be applied to its own output, we can define \emph{dream path}:

\begin{eqnarray}
\Pi(b,g)&\defeq&c_0,\ldots, c_n\in\mathit{Chunks}\{n\}\\
c_i&\defeq&\begin{cases}
\mathcal{G}\idx{4K}(g) &\text{if }i=0\\   
\Delta(b,c_{i-1}) &\text{otherwise}
\end{cases}
\end{eqnarray}

The dream path function is a pseudo random generator defined by finite recursion. A \emph{dream} is a particular pairing of a generator and a one time pad that are input and output of a dream path.
Given a dream path of length $n$ and grantee overlay address $a$ and a batch $b$ with uniformity depth $d$, uploader needs to find the generator $g$, such that the $n$-th chunk of the dream path should land in $a$'s neighbourhood (i.e., $\mathit{PO}(a,\textsc{address}(c_{n-1}))\geq d$).

\begin{eqnarray}
\mathit{dream}(b,n,a) &\subset &\mathit{Keys}\times\mathit{Chunks}\\
\langle g, k\rangle &\in& \mathit{dream}(b,n,a) \\
&&\Longleftrightarrow\\
&&k=\Pi(b,g)\idx{n}
\land\\
&&\mathit{PO}(a,H_{\mathit{bmt}}(k))\geq d
\end{eqnarray}

The uploader is able to construct the dream pad.
As a result they can also calculate $k$ given $g$. Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$.

Now we can turn to the definition of a dream, a network protocol based access method that is deniable, revocable, expirable, addressable, malleable. In order for downloaders to calculate the dream pad, they must rely on the network, whereby each recursive step is calculated by a node in the neighbourhood designated by the respective input chunk's address.

In order to guarantee the correct termination, the following criteria must be fulfilled:
\begin{itemize}[noitemsep]
\item[---] all  buckets of batch $b$ designated by $p_0,\ldots, p_{n-1}$ amust be non-empty
\item[---] uniformity depth of the batch must remain higher than the storage depth of nodes in the neighbourhoods designated by $p_0, \ldots , p_{n-1}$.
\item[---] the output of each step produced by a node in the neighbourhood designated by $p_i$ must be sent to the 
neighbourhood designated by $p_{i+1}$. 
\item[---]  nodes in the neighbourhoods on each hop of the dream path (designated by $p_0,\ldots, p_{n-1}$) must be 
incentivised to compute the OTP update as well as send off the resulting output chunk to the next neighbourhood. \end{itemize}

\begin{figure}[htbp]
  \centering
    % \includegraphics[width=0.8\textwidth]{}
  \caption{The dream protocol}
\label{fig:protocol}
\end{figure}

Let us  define a network protocol \emph{dream} (see figure \ref{fig:protocol}). Those who play the protocol, listen to single owner chunks wrapping the input.
We assume that they extract the batch identifier $b$, and the payload CAC $c$ as parameters to the OTP update function and then calculate the output pad, wrap it as a dream chunk and send it to the network towards the address of the output chunk so that the destination neighbourhood 
can calculate the next step. If the protocol is followed up to $n$ steps, then $k=c(n)$ is received by the target node at address $a$.

We now turn to the construction of dream chunks. 
Uploader $U$ wishes to grant downloader $D$ (at overlay address $a$) revocable  access to content chunk $C$.
$U$  chooses a postage batch $b$ it owns that is completely filled, a step-count $n$, and a depth $d$.
$U$ creates a dream generator $\langle g, k\rangle\in\mathit{dream}(b,d,n,a)$.

$U$ calculates the "ciphertext" $C'=C \xor k$ and uploads it to swarm obtaining $r=\mathit{H}(C')$ as parity reference. Now $U$ creates the \emph{dream reference} as $\mathit{ref}(C)=\langle r,b,g\rangle $ which can be given to the downloader grantee $D$.%
%
\footnote{So far, references to swarm content included an address (\emph{Swarm hash} for unencrypted content) or an address with a decryption key (for encrypted content). The dream reference now defines a third type of reference, serialised in four segments comprising the parity address, the batch id, the initial generator and a decryption key.}

\subsection{Analysis}

% \subsection{Retrieval}
% \label{def:retrieval}
$D$ in possession of a dream chunk reference as 
$\langle r,b,g\rangle $ calculated by $U$  as $\mathit{ref}(C)$ constructs $p=\mathcal{G}\idx{4K}(g)$ and wraps it as a dream chunk and sends it to swarm. When it receives the dream chunk for $b$, it extracts payload $k$.

$D$ retrieves the parity chunk $C'$ using reference $r$ and then decodes $C=C'\xor k$.
The retrieval process is trivially correct as long as 
\begin{enumerate}[noitemsep]
\item[---]  The parity chunk $C'$ is retrievable via normal retrieval. 
\item[---] The dream protocol is adhered to by cooperating nodes.
\item[---] The batch bucket bottom has not changed for the neighbourhoods that are on the dream path. 
\end{enumerate}

We are now turning to access revocation. Deletion of content is equivalent to revoking all access.

Uploader $U$ had granted $D$ (at address $a$) access to $C$ through the dream reference $\langle r,b,g\rangle $. $U$ revokes access from $D$ by uploading extra chunks to batch $b$ such that $\overline{\Huge\chi}$ changes.

As long as there is a single honest neighbourhood walked by the protocol, such that the nearest chunk to pivot $p$ from batch $b$ changes, then $D$ can no longer retrieve $C$.

To prove this, first let us assume that there is a bucket designated by some $p_i$ on the dream path such that the $\overline{\Huge\chi}$ for the bucket changed since uploader constructed the dream. When the participating node in the neighbourhood designated by $p_i$ calculates the OTP update function, the result will be completely different and the dream chain will not terminate. \qedsymbol

% \subsection{Incentives}

% Discuss the reliability of the protocol in the context of economic incentives.
% \begin{figure}[htbp]
%   \centering
%     \includegraphics[width=0.8\textwidth]{figs/push-payment.pdf}
%   \caption{Incentives for push syncing.}
% \label{fig:push-payment}
% \end{figure}

% How to determine when to forward.
% discuss the cost of a dream

\subsection{Security}

The reference does not leak either the step-count or any other information about the neighbourhoods touched. Neither do the participants in the protocol know which position they are at in the chain or anything  about the other neighbourhoods except for the following one to which they are push-syncing their chunk.

The construction is deniable.
Beside our sensitive content $C$, take $A$, any uncontroversial content chunk. When producing $C'$, the owner also produces  $A'=A\xor k$ and upload it. When asked about $k$, producing $A$ makes the denial of other content including $C$ more plausible. \qedsymbol

% Swarm is a decentralized network where individual participants cannot (and should not) be trusted to play by the rules; it is the built-in incentives that encourage cooperative behavior and help maintain it, it being a Nash-equilibrium. The condition for content deletion to be achieved is that in \emph{at least one neighborhood} all nodes are honest, i.e., do not operate a malicious node which would store and serve information that they should not store and serve.% 
% %
% \footnote{With the growth of the Swarm network, the number of neighborhoods increases whereas the membership of neighborhoods remains about the same. Thus, the costs of attacking the proposed scheme are expected to increase with network size.}
% %

A powerful adversary can infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it surely) and keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand. But there is a huge price on such indiscriminate archiving of all Swarm's content and that is the only way to reliably defeat the dream construction.

We now turn to the discussion how to calibrate the step-count given a security model. Let us assume a network-wide neighbourhood infiltration rate of $\frac{1}{2}$, meaning that  half of the neighbourhoods in the network is assumed  to be  malicious and colluding.

Assuming that when  revoking access, the owner uploaded new chunks in each neighbourhood on the dream path. 
Given a particular dream chain, 
However, if even  a single neighbourhood on the dream path is honest they respect the newly arrived chunks and divert the dream path, making it never terminate with the grantee.

% ZWe assume that malicious nodes are able and willing to disregard these  new chunks and in an attempt to facilitate abuse of revoked access respond to the ream protocol just like  before the revocation. if each  neighbourhood  on  the path is malicious, it is in principle possible that the dream protocol gives the unchanged response, and access is abused, i.e., allowed even though it is revoked.
Thus for  a breach of access to happen, all neighbourhoods must be malicious.
In our security model a neighbourhood is malicious with  uniform and independent probability.  For an overall infiltration rate of 1 out of $k$, the chance of all neighbourhoods on a given random dream path being malicious is $k^{-n}$. For a security requirement of success rate of $\sigma$ "nines", i.e error rate less than $10^{-\sigma}$,  we can formulate the requirement as

\begin{equation}
    \frac{1}{k^n}< \frac{1}{10^{-\sigma}}
\end{equation}
Now, expressing $k$ as $10^\kappa$, taking the logarithm of both sides and multiply by $-1$, we get

\begin{equation}
    n > \frac{\sigma}{\kappa}
\end{equation}

With one in every 10 neighbourhoods being malicious, the dream path must have as many hops as the number of nines expressing the success rate.%
%
\footnote{I.e., with malicious node probability $0.1$, a 6-hop-long path will effectively revoke access wtith certainty of $99.9999\%$.}

% Since these parts are symmetrical, it is somewhat misleading to refer to them as ciphertext and key.  However, in the deletable construct one part will be available via a normal hash reference, this we can call one time pad, while the ephemeral part is called encoded content.

% There is also a variation that "self-destructs", i.e. that deletes itself upon
% retrieval, but it is important to understand that it is a very fragile
% mechanism with no guarantee that the intended recipient will be able to
% download it even once. It should only be used in settings where the recipient can,
% for a limited time, request re-uploading in case the content gets deleted
% without all its parts reaching the recipient.

% \begin{definition}[pss encode function]

% \begin{equation}
% \Psi(b,d,c)=\mathit{Trojan}(\mathit{nonce}, \mathit{topic}, \mathit{payload})
% \end{equation}
% where
% \begin{equation}
% \mathit{nonce} = c[0:32]
% \end{equation}
% \begin{equation}
% \mathit{topic} = \tau
% \end{equation}
% \begin{equation}
% \mathit{payload}[0:32] =  b
% \end{equation}
% \begin{equation}
% \mathit{payload}[32:34] = \mathit{bigendian.Uint16}(d)
% \end{equation}
% \begin{equation}
% \mathit{payload}[34:4030] = c[100:4096]
% \end{equation}
% \end{definition}

### 2024-06-11 dream.tex
Updates from Overleaf

@@
 
-The increasing trend of digitisation and the ensuing ubiquity of personal data in the cloud touch on sensitivities more than ever before. 
-In response to the former, regulators often formulate requirements on removability that fails to recognise 
-that it is virtually impossible to make material once seen become unseen.
-Entities that operate infrastructure underpinning digital publishing are in control of the technology serving the content and  are able to implement "removal" by denying access.
-Such centralised gate-keeping  has its regulatory appeal as it can be thought of as an instrument of law enforcement.
-Although such intervention rarely solves the breach of unrealistic privacy rights, the gate-keeper provides a well-defined responsible party with the effective course of action: not serving censored content.
-While this gives the unwary a phony sense of security, it only exemplifies the bigger problem of censorship: publishing platforms have the technological means to filter content and centralised control makes exerting influence on content cheaper.
-What starts out as sensible measures of content curation with time becomes extant censorship.
-This is all the more problematic with social platforms that gained quasi-monopolistic status due to netwtork effects.
-As a result of high switching cost, innocent content creators ever more often end up being 'deplatformed'.
-Similarly, the ability to identify hosts and deny access to content via legal decree gives powers to be the means to infringe on freedom of speech.
+The increasing digitisation of data and the widespread use of cloud services have amplified concerns surrounding the privacy and control of personal information more than ever before. 
+In response, regulators often impose requirements on removability without recognising that it is virtually impossible to make material once seen become unseen.
+Entities that operate infrastructure underpinning digital publishing are in control of the technology serving the content, making it possible for them to carry out "removal" by simply denying access.
+Such centralised gatekeeping may seem appealing from a regulatory standpoint, as it can be seen as an instrument of law enforcement.
+Although such interventions rarely solve the breach of unrealistic privacy rights, the gatekeeper still provides a well-defined responsible party with the effective course of action: not serving censored content.
+While this gives the unwary a false sense of security, it only exemplifies the larger issue of censorship: publishing platforms possess the technological means to filter content, and centralised control makes it more affordable to exert influence over the content.
+What starts out as sensible measures of content curation gradually transforms into extant censorship.
+This is all the more problematic with social platforms that have gained quasi-monopolistic status as a result of network effects.
+Due to the high costs associated with switching platforms, innocent content creators increasingly find themselves 'deplatformed'.
+Similarly, the ability to identify hosts and deny access to content through legal means gives powers to be the means to infringe on freedom of speech.
 
-In the decentralised paradigm of web 3, there is no longer an operating entity behind either the publishing platform or the hosting infrastructure.
-Therefore, the cost of censorship is raised beyond the realm of feasibility.
-And yet, the potential permanent exposure of personal data is perceived as scary to most and some novel solution is called for to restore the sense of security gate-keepers purportedly represent.
+In the decentralised paradigm of web 3, there is no longer a single operating entity controlling the publishing platform or the hosting infrastructure.
+This renders the the cost of censorship infeasible.
+And yet, the potential permanent exposure of personal data remains a significant concern for most individuals.
+Therefore, there is a need for novel solutions that can restore the sense of security that centralised gatekeepers purportedly represent.
 
@@
 
-First, notice that any reference to removal of information in the sense of erasing it from all physical storage devices is both unenforceable and impractical.
-Even the strictest data protection audits do not require the erasure of offending data from backup tapes.
-In general, the tacit assumption is that information ordered to be removed should become inaccessible through \emph{typical precedented methods of access}.
+First, note that any reference to removal of information in the sense of erasing it from all physical storage devices is both unenforceable and impractical.
+Even the most rigorous data protection audits do not require the erasure of offending data from backup tapes.
+In general, the tacit assumption is that information ordered to be removed should become inaccessible through \emph{typical, precedented methods of access}.
 
-In what follows we formulate the strongest meaningful definition of deletion applicable to decentralised storage and offer a construction that implements it.
-Importantly, such a construction is purely technical, referring to primary actors' capabilities and costs as opposed to procedural, referring to intermediaries' obligations to respect the rights of primary actors.
+In what follows, we formulate the strongest meaningful definition of deletion applicable to decentralised storage systems and offer a construction that implements it.
+Importantly, this approach is purely technical, focusing on thecapabilities and costs of primary actors, rather than relying on  procedural measures that impose obligations on intermediaries to respect the rights of primary actors.
 
-Primary actors here are the \emph{uploader} that wishes to share content by granting read access to a number of parties, called \emph{downloaders}.
-Granting access is defined as providing some canonical \emph{reference} to the content that the system can eventually resolve to the full information meant to be disclosed.
-Any party that is privileged to access this information, is able to store, re-code and potentially disseminate it.
-This provides a myriad of ways to make accessible the content at any later point in time to get round whatever process would qualify as deletion or removing access.
+The primary actors in this context are the \emph{uploaders} who wish to share content by granting read access to a number of parties, called \emph{downloaders}.
+Granting access is defined as providing a canonical \emph{reference} to the content, allowing the system to eventually retrieve to the complete information that is intended to be disclosed.
+Any party that is privileged to access this information is able to store, re-code, and potentially disseminate it.
+This provides a myriad of ways to make the content accessible at any later point in time, bypassing any process that would qualify as deletion or removal of access.
 There is, by definition, no protection possible against such adversity.
-As a consequence, any legally sound notion of deletion (retrospective revocation of access) is meant to be interpreted in a narrower sense: \emph{the viability to replay the same access method}  for a cost lower than the \emph{full cost} of storing the content.%
+As a consequence, any legally sound notion of deletion (retrospective revocation of access) is meant to be interpreted in a narrower sense: \emph{the viability to replay the same access method}  at a lower cost than the \emph{full cost} of storing the content.%
 %
-\footnote{That is, the total cost of storage paid for the full size for the entire period kstarting from the time access was revoked uptill the attempted breach.}
+\footnote{That is, the total storage cost paid for the full size of the content starting  starting from the time access was revoked up till the attempted breach.}
 
-Let us now define deletion as a scheme for uploading content with access revocation that has the following requirements:
+Let us now define deletion as a scheme for uploading content with access revocation, meeting the following requirements:
 
 \begin{itemize}[noitemsep]
-    \item[--] \emph{specialisation} -- Uploader is able to choose at the time of publishing a specialised construct%
+    \item[--] \emph{specialisation} -- The uploader is able to choose at the time of publishing a specialised construct%
 %
 \footnote{From a user's perspective, content that is meant to be reliably deleted should be uploaded as such.
-The costs of uploading such content is allowed to be a (small integer) multiple of the cost of regular, censorship-resistant but non-deletable uploads.}
+The cost of uploading such content is allowed to be a (small integer) multiple of the cost of regular, censorship-resistant but non-deletable uploads.}
 %
     allowing retrospective revocation of access.
-    \item[--] \emph{sovereignty} -- Uploader is the sovereign owner of the data and is in sole possession of the means to selectively revoke access from any party previously granted access.%
+    \item[--] \emph{sovereignty} -- The uploader is the sovereign owner of the data and is in sole possession of the means to selectively revoke access from any party that was previously granted access.%
 %
-\footnote{The uploader's credentials are necessary to delete their own deletable content. Deletables must also allow access control, i.e., only available to a specific set of recipients.}
+\footnote{The uploader's credentials are necessary to delete their own deletable content. Deletables must also allow access control, i.e., they are only available to a specific set of recipients.}
 %
-    \item[--] \emph{security} -- After revocation of access, a grantee is unable to access the content using the same reference that the access was granted with, or any other cue shorter than the deleted content.%
+    \item[--] \emph{security} -- After the revocation of access, a grantee is unable to access the content using the same reference or any other cue shorter than the deleted content.%
 %
@@
 
-Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and it is not accessed sufficiently often, the chunks constituting it will be garbage collected.
-However, chunks with  expired postage stamps cannot be regarded as reliably deleted to satisfy the requirements of sovereignty and security.
+Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and the content is not accessed frequently, the chunks constituting it will be garbage-collected.
+However, chunks with expired postage stamps cannot be regarded as reliably deleted in order to to satisfy the requirements of sovereignty and security.
 
@@
 The goal of this section is to arrive at a formal construction of a DISC-based revocable access model.
-We will restrict our scheme to \emph{chunks}, the fundamental fix sized storage unit of Swarm's DISC model.
-The proposed \emph{dream} construct implements a  deletable content storage and access model conforming to the requirements of specialisation, sovereignty, and security.
+We will restrict our scheme to \emph{chunks}, the fundamental fixed-size storage units of Swarm's DISC model.
+The proposed \emph{dream} construct implements a deletable content storage and access model that fulfills the requirements of specialisation, sovereignty, and security.
 The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model.
-On top of this, as customary in swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} expressing requirements of access control.
+On top of this, as is customary in Swarm, it serves as a mnemonic acronym, resolving to the 5 \emph{dream attributes} expressing requirements of access control.
 
 \begin{itemize}
-    \item[\textbf{D}] \emph{deniable} -- the dream key serves as a one time pad for decryption.
-Since multiple content chunks (in fact any arbitrary content) can use the same dream pad, the key's association to any content is plausibly deniable.
-    \item[\textbf{R}] \emph{revocable} -- access granted by dream keys is revocable. Revoking access from all parties including oneself is taken as deletion.   
-    \item[\textbf{E}] \emph{expirable} -- the scheme allows for one-time use, i.e., the key can only be retrieved once.
-    \item[\textbf{A}] \emph{addressable} -- access can be granted to a neighbourhood, only clients operating a node with an overlay address in a given range  are able to access.
-    \item[\textbf{M}] \emph{malleable}  --  the construct is resilient to churn and dynamic change of network size, is reusable across independent grantees and upgradeable.
+    \item[\textbf{D}] \emph{deniable} -- The dream key serves as a one-time pad for decryption.
+Since multiple content chunks (in fact, any arbitrary content) can use the same dream pad, the key's association to any content is plausibly deniable.
+    \item[\textbf{R}] \emph{revocable} -- Access granted through dream keys is revocable. Revoking access from all parties, including oneself, is considered deletion.   
+    \item[\textbf{E}] \emph{expirable} -- The scheme allows for one-time use, i.e., the key can only be retrieved once.
+    \item[\textbf{A}] \emph{addressable} -- Access can be granted to a neighbourhood, where only clients operating a node within a particular overlay address range are able to access the content.
+    \item[\textbf{M}] \emph{malleable}  --  The construct is resilient to churn and dynamic changes in network size; it is reusable across independent grantees and upgradeable.
 \end{itemize}
 
-The scheme is built on top of DISC's APIs and can be implemented entirely as a second layer solution.
-Despite its rich feature set, the scheme does not use complex cryptography, but relies on the interplay of various component subsystems.
+The scheme is built on top of DISC's APIs and can be implemented entirely as a second-layer solution.
+Despite its rich feature set, the scheme does not use complex cryptography, but instead leverages the interplay of various component subsystems.
 
@@
 
-Assume we have a pseudo-random deterministic function that generates a longer (e.g., chunk-sized = 4K) sequence $c$ from a key-sized (32 byte) generator $g$.%
+Assume we have a pseudo-random deterministic function that generates a longer (e.g., chunk-sized = 4K) sequence $c$ from a key-sized (32-byte) generator $g$.%
 %
 \footnote{For example, the Keccak sponge function used throughout Ethereum for hashing does have this capability.
-Alternatively, the blockcipher encryption using initial nonce $g$ could be applied to a  fixed constant chunk such as all zeros.}
+Alternatively, the blockcipher encryption using the initial nonce $g$ could be applied to a fixed constant chunk, such as all zeros.}
 
@@
 
-The central construct called a \emph{dream} is a chunk-sized one time pad which acts as a decryption key for the deletable content.
+The central construct called a \emph{dream} is a chunk-sized one-time pad which acts as a decryption key for the deletable content.
 The dream chunk is collectively created by a set of nodes following a network protocol.
-Each node in the sequence of participating peers receives a piece of data and uses it as input together with data present in their reserve to produce an output sent to the neighbourhood of the next node in the chain.
-As long as each node in the chain performs the same calculation and forwards the same result, the same one time pad can be construed and ultimately the deletable content can be retrieved and read.
+Each node in the sequence of participating peers receives a piece of data and combines it as input with data from their reserve to produce an output, which is then sent to the neighbourhood of the next node in the chain.
+As long as each node in the chain performs the same calculation and forwards the same result, the same one-time pad can be generated, enabling the retrieval and reading of the deletable content.
 
-The insight here is that retrieving a deletable chunk's references involves calculations using a set of immutable chunks under the control of the uploader.
+The insight here is that retrieving a deletable chunk's references involves calculations using a set of immutable chunks controlled by the uploader.
 The uploader's ability to change this underlying set holds the key to providing the necessary mutability required by any notion of deletion.
 
-Let $b$ be the batch ID (a 32-byte hash) of a postage batch owned by $U$ and 
+Let $b$ be the batch ID (a 32-byte hash) of a postage batch owned by $U$, and 
 let $\mathit{Chunks}(\gamma,b,p)$ stand for all chunks stamped by $b$ at block $\gamma$ in the bucket designated by the pivot $p$.
@@
 may not contain any chunks belonging to the batch bucket.
-In such cases there is no chunk closest to $p$, thus the \emph{BBB} function is undefined.
+As a result, there will not be a chunk closest to $p$, rendering the \emph{BBB} function undefined in such cases.
 
@@
 
-Since uniformity depth of batches is meant to be larger than nodes' storage depth, all chunks in a batch bucket designated with $p$ is  in the reserve of every node in the neighbourhood designated by $p$.
+Since the uniformity depth of batches is intended to be greater than the nodes' storage depth, it is expected that all chunks in a batch bucket designated with $p$ are present in the reserve of every node in the neighbourhood designated by $p$.
 
-If swarm grows and  there are nodes whose storage  depth is higher than the batch's uniformity depth, then the neighbourhood designated by $p$ may not contain all the
+However, if Swarm grows and there are nodes whose storage  depth is higher than the uniformity depth of the batch, then the neighbourhood designated by $p$ may not contain all the
 chunks in the bucket, thus compromising the computability of the OTP update function.
 
-As long as 1) the bucket of batch $b$ designated by $p$ is not empty and 2) the neighbourhoods include (one or more) buckets then the update function is computable by any storer node within the neighbourhood designated by $p$.
+Nevertheless, if two conditions are met, namely: 1) the bucket of batch $b$ designated by $p$ is not empty and 2) the neighbourhoods include one or more buckets, then the update function is computable by any storer node within the neighbourhood designated by $p$.
 
-Since the task function can be applied to its own output, we can define \emph{dream path}:
+Since the update function can be applied to its own output, we can define the \emph{dream path} as follows:
 
@@
 
-The dream path function is a pseudo random generator defined by finite recursion. A \emph{dream} is a particular pairing of a generator and a one time pad that are input and output of a dream path.
-Given a dream path of length $n$ and grantee overlay address $a$ and a batch $b$ with uniformity depth $d$, uploader needs to find the generator $g$, such that the $n$-th chunk of the dream path should land in $a$'s neighbourhood (i.e., $\mathit{PO}(a,\textsc{address}(c_{n-1}))\geq d$).
+The dream path function is a pseudo-random generator that is defined by finite recursion.
+A \emph{dream} is a particular pairing of a generator and a one-time pad, which serve as the input and output of a dream path.
+Given a dream path of length $n$, a grantee overlay address $a$, and a batch $b$ with a uniformity depth of $d$, the uploader needs to find the generator $g$ such that the $n$-th chunk of the dream path falls within $a$'s neighbourhood (i.e., $\mathit{PO}(a,\textsc{address}(c_{n-1}))\geq d$).
 
@@
 
-The uploader is able to construct the dream pad.
-As a result they can also calculate $k$ given $g$. Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$.
+The uploader is able to construct the dream pad, which also allows them to calculate $k$ given $g$.
+Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$.
 
-Now we can turn to the definition of a dream, a network protocol based access method that is deniable, revocable, expirable, addressable, malleable.
-In order for downloaders to calculate the dream pad, they must rely on the network, whereby each recursive step is calculated by a node in the neighbourhood designated by the respective input chunk's address.
+Now we can turn to the definition of a dream, which is a network protocol-based access method that is deniable, revocable, expirable, addressable, and malleable.
+In order for downloaders to calculate the dream pad, they must rely on the network, where each recursive step is calculated by a node in the neighbourhood designated by the respective input chunk's address.
 
@@
 \begin{itemize}[noitemsep]
-\item[---] all  buckets of batch $b$ designated by $p_0,\ldots, p_{n-1}$ amust be non-empty
-\item[---] uniformity depth of the batch must remain higher than the storage depth of nodes in the neighbourhoods designated by $p_0, \ldots , p_{n-1}$.
-\item[---] the output of each step produced by a node in the neighbourhood designated by $p_i$ must be sent to the 
-neighbourhood designated by $p_{i+1}$. 
-\item[---]  nodes in the neighbourhoods on each hop of the dream path (designated by $p_0,\ldots, p_{n-1}$) must be 
-incentivised to compute the OTP update as well as send off the resulting output chunk to the next neighbourhood. \end{itemize}
+\item[---] all  buckets of batch $b$ designated by $p_0,\ldots, p_{n-1}$ must be non-empty
+\item[---] the uniformity depth of the batch must remain higher than the storage depth of nodes in the neighbourhoods designated by $p_0, \ldots , p_{n-1}$
+\item[---] the output of each step, produced by a node in the neighbourhood designated by $p_i$, must be sent to the 
+neighbourhood designated by $p_{i+1}$ 
+\item[---]  nodes in the neighbourhoods along each hop of the dream path (designated by $p_0,\ldots, p_{n-1}$) must be 
+incentivised to compute the OTP update and forward the resulting output chunk to the next neighbourhood. \end{itemize}
 
@@
 
-Let us  define a network protocol \emph{dream} (see figure \ref{fig:protocol}). Those who play the protocol, listen to single owner chunks wrapping the input.
-We assume that they extract the batch identifier $b$, and the payload CAC $c$ as parameters to the OTP update function and then calculate the output pad, wrap it as a dream chunk and send it to the network towards the address of the output chunk so that the destination neighbourhood
-can calculate the next step. If the protocol is followed up to $n$ steps, then $k=c(n)$ is received by the target node at address $a$.
+Let us  define a network protocol \emph{dream} (see Figure \ref{fig:protocol}). Participants in the protocol listen to single-owner chunks that wrap the input.
+We assume that they extract the batch identifier $b$ and the payload CAC $c$ as parameters to the OTP update function.
+Then they calculate the output pad, wrap it as a dream chunk, and send it to the network towards the address of the output chunk.
+This allows the destination neighbourhood
+to calculate the next step. If the protocol is followed up to $n$ steps, then the target node receives $k=c(n)$ at address $a$.
 
@@
 Uploader $U$ wishes to grant downloader $D$ (at overlay address $a$) revocable  access to content chunk $C$.
-$U$  chooses a postage batch $b$ it owns that is completely filled, a step-count $n$, and a depth $d$.
-$U$ creates a dream generator $\langle g, k\rangle\in\mathit{dream}(b,d,n,a)$.
-
-$U$ calculates the "ciphertext" $C'=C \xor k$ and uploads it to swarm obtaining $r=\mathit{H}(C')$ as parity reference.
-Now $U$ creates the \emph{dream reference} as $\mathit{ref}(C)=\langle r,b,g\rangle $ which can be given to the downloader grantee $D$.%
+$U$  chooses a postage batch $b$ it owns, ensuring that it is not completely filled. Additionally, $U$ chooses a step-count $n$ and a depth $d$.
+$U$ creates a dream generator $\langle g, k\rangle\in\mathit{dream}(b,d,n,a)$, then calculates the "ciphertext" $C'=C \xor k$ and uploads it to Swarm, obtaining $r=\mathit{H}(C')$ as the parity reference.
+Now $U$ creates the \emph{dream reference} as $\mathit{ref}(C)=\langle r,b,g\rangle $, which can be given to the downloader grantee $D$.%
 %
 \footnote{So far, references to swarm content included an address (\emph{Swarm hash} for unencrypted content) or an address with a decryption key (for encrypted content).
-The dream reference now defines a third type of reference, serialised in four segments comprising the parity address, the batch id, the initial generator and a decryption key.}
+The dream reference now defines a third type of reference, serialised in four segments comprising the parity address, the batch ID, the initial generator, and a decryption key.}
 
@@
 % \label{def:retrieval}
-$D$ in possession of a dream chunk reference as 
-$\langle r,b,g\rangle $ calculated by $U$  as $\mathit{ref}(C)$ constructs $p=\mathcal{G}\idx{4K}(g)$ and wraps it as a dream chunk and sends it to swarm.
-When it receives the dream chunk for $b$, it extracts payload $k$.
+$D$, in possession of a dream chunk reference as 
+$\langle r,b,g\rangle $ calculated by $U$  as $\mathit{ref}(C)$, constructs $p=\mathcal{G}\idx{4K}(g)$, wraps it as a dream chunk, and sends it to Swarm.
+When $D$ receives the dream chunk for $b$, it extracts payload $k$.
 
 $D$ retrieves the parity chunk $C'$ using reference $r$ and then decodes $C=C'\xor k$.
-The retrieval process is trivially correct as long as 
+The retrieval process is trivially correct as long as: 
 \begin{enumerate}[noitemsep]
-\item[---]  The parity chunk $C'$ is retrievable via normal retrieval. 
-\item[---] The dream protocol is adhered to by cooperating nodes.
-\item[---] The batch bucket bottom has not changed for the neighbourhoods that are on the dream path. 
+\item[---]  The parity chunk $C'$ is retrievable via normal retrieval methods. 
+\item[---] The dream protocol is followed by cooperating nodes.
+\item[---] The batch bucket bottom remains unchanged for the neighbourhoods along the dream path. 
 \end{enumerate}
 
-We are now turning to access revocation. Deletion of content is equivalent to revoking all access.
+We are now turning to access revocation, which is equivalent to deleting content by revoking all access.
 
 Uploader $U$ had granted $D$ (at address $a$) access to $C$ through the dream reference $\langle r,b,g\rangle $.
-$U$ revokes access from $D$ by uploading extra chunks to batch $b$ such that $\overline{\Huge\chi}$ changes.
+$U$ revokes $D$'s access by uploading extra chunks to batch $b$ such that $\overline{\Huge\chi}$ changes.
 
-As long as there is a single honest neighbourhood walked by the protocol, such that the nearest chunk to pivot $p$ from batch $b$ changes, then $D$ can no longer retrieve $C$.
+As long as there is at least one honest neighbourhood walked by the protocol, such that the nearest chunk to pivot $p$ from batch $b$ changes, downloader $D$ will no longer be able to retrieve content $C$.
 
-To prove this, first let us assume that there is a bucket designated by some $p_i$ on the dream path such that the $\overline{\Huge\chi}$ for the bucket changed since uploader constructed the dream.
+To prove this, first let us assume that there is a bucket designated by some $p_i$ on the dream path such that the $\overline{\Huge\chi}$ for the bucket has changed since the uploader constructed the dream.
 When the participating node in the neighbourhood designated by $p_i$ calculates the OTP update function, the result will be completely different and the dream chain will not terminate.
-\qedsymbol
 
@@
 
-The reference does not leak either the step-count or any other information about the neighbourhoods touched.
-Neither do the participants in the protocol know which position they are at in the chain or anything  about the other neighbourhoods except for the following one to which they are push-syncing their chunk.
+The reference does not leak any information about the step-count or the neighbourhoods involved.
+Neither are the participants in the protocol aware of their position in the chain or any detailsabout the other neighbourhoods, except for the immediate next one to which they are push-syncing their chunk.
 
-The construction is deniable.
-Beside our sensitive content $C$, take $A$, any uncontroversial content chunk.
-When producing $C'$, the owner also produces  $A'=A\xor k$ and upload it.
-When asked about $k$, producing $A$ makes the denial of other content including $C$ more plausible.
-\qedsymbol
+The construction of the dream reference is deniable.
+Beside our sensitive content $C$, let us consider $A$, any uncontroversial content chunk.
+When producing $C'$, the owner also produces  $A'=A\xor k$ and uploads it.
+When asked about $k$, producing $A$ makes the denial of other content, including $C$, more plausible.
 
@@
 
-A powerful adversary can infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it surely) and keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand.
-But there is a huge price on such indiscriminate archiving of all Swarm's content and that is the only way to reliably defeat the dream construction.
+A powerful adversary could potentially infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it).
+They could also keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand.
+However, there is a huge cost associated with such indiscriminate archiving of all Swarm's content, and that is the only way to reliably defeat the dream construction.
 
-We now turn to the discussion how to calibrate the step-count given a security model.
-Let us assume a network-wide neighbourhood infiltration rate of $\frac{1}{2}$, meaning that  half of the neighbourhoods in the network is assumed  to be  malicious and colluding.
+We now turn to the discussion of how to calibrate the step-count in relation to the security model.
+Let us assume a network-wide neighbourhood infiltration rate of $\frac{1}{2}$, meaning that  half of the neighbourhoods in the network are assumed  to be  malicious and colluding.
 
-Assuming that when  revoking access, the owner uploaded new chunks in each neighbourhood on the dream path. 
-Given a particular dream chain, 
-However, if even  a single neighbourhood on the dream path is honest they respect the newly arrived chunks and divert the dream path, making it never terminate with the grantee.
+When access is revoked, the owner uploads new chunks in each neighbourhood along the dream path. 
+Given a particular dream chain, however, if even a single neighbourhood on the dream path is honest, they will respect the newly arrived chunks and divert the dream path, preventing it from terminating with the grantee.
 
@@
 if each  neighbourhood  on  the path is malicious, it is in principle possible that the dream protocol gives the unchanged response, and access is abused, i.e., allowed even though it is revoked.
-Thus for  a breach of access to happen, all neighbourhoods must be malicious.
-In our security model a neighbourhood is malicious with  uniform and independent probability.
+Thus, for  a breach of access to happen, all neighbourhoods must be malicious.
+In our security model, a neighbourhood is malicious with  uniform and independent probability.
 For an overall infiltration rate of 1 out of $k$, the chance of all neighbourhoods on a given random dream path being malicious is $k^{-n}$.
-For a security requirement of success rate of $\sigma$ "nines", i.e error rate less than $10^{-\sigma}$,  we can formulate the requirement as
+For a security requirement of success rate of $\sigma$ "nines", where the error rate is less than $10^{-\sigma}$, we can formulate the requirement as
 
@@
 \end{equation}
-Now, expressing $k$ as $10^\kappa$, taking the logarithm of both sides and multiply by $-1$, we get
+Now, expressing $k$ as $10^\kappa$, taking the logarithm of both sides and multiplying by $-1$, we get
 
@@
 %
-\footnote{I.e., with malicious node probability $0.1$, a 6-hop-long path will effectively revoke access wtith certainty of $99.9999\%$.}
+\footnote{I.e., with malicious node probability $0.1$, a 6-hop-long path will effectively revoke access with a certainty of $99.9999\%$.}
 

### 2025-09-11 archived/dream.tex
revision submitted to acm  tweb special issue

--- 2024-06-11 dream.tex
@@
-\section{Dream}
+\section{Introduction}
 
@@
 In the decentralised paradigm of web 3, there is no longer a single operating entity controlling the publishing platform or the hosting infrastructure.
-This renders the the cost of censorship infeasible.
+This renders the cost of censorship infeasible.
 And yet, the potential permanent exposure of personal data remains a significant concern for most individuals.
@@
 
-\subsection{Deletion and revoking access}
+\section{Deletion and revoking access}
 
@@
 In what follows, we formulate the strongest meaningful definition of deletion applicable to decentralised storage systems and offer a construction that implements it.
-Importantly, this approach is purely technical, focusing on thecapabilities and costs of primary actors, rather than relying on  procedural measures that impose obligations on intermediaries to respect the rights of primary actors.
+Importantly, this approach is purely technical, focusing on the capabilities and costs of primary actors, rather than relying on  procedural measures that impose obligations on intermediaries to respect the rights of primary actors.
 
@@
 Note that normally uploaded content may be \emph{forgotten}: if nobody pays for storing it and the content is not accessed frequently, the chunks constituting it will be garbage-collected.
-However, chunks with expired postage stamps cannot be regarded as reliably deleted in order to to satisfy the requirements of sovereignty and security.
+However, chunks with expired postage stamps cannot be regarded as reliably deleted in order to satisfy the requirements of sovereignty and security.
 
-\subsection{Construction}
+\section{Construction}
 
@@
 We will restrict our scheme to \emph{chunks}, the fundamental fixed-size storage units of Swarm's DISC model.
-The proposed \emph{dream} construct implements a deletable content storage and access model that fulfills the requirements of specialisation, sovereignty, and security.
+The proposed \emph{dream} construct implements a deletable content storage and access model that fulfils the requirements of specialisation, sovereignty, and security.
 The use of the word 'dream' alludes to the somewhat unexpected finding that such a construct is even possible in the immutable DISC model.
@@
 
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.65\textwidth]{figs/dream.pdf}
+  \caption{The dream protocoll: collective decryption of content}
+\label{fig:disc}
+\end{figure}
+
 The dream path function is a pseudo-random generator that is defined by finite recursion.
@@
 The uploader is able to construct the dream pad, which also allows them to calculate $k$ given $g$.
-Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$.
+Since $\mathit{H}(k)$ is uniform, the chance of $\mathit{PO}(a,\mathit{H}(k))\geq d$ is 1 in $2^{d}$, therefore the uploader can find a suitable generator $g$ in $2^d$ attempts on average.
 
@@
 
-\subsection{Analysis}
+\section{Analysis}
 
@@
 
-\subsection{Security}
+\section{Security}
 
 The reference does not leak any information about the step-count or the neighbourhoods involved.
-Neither are the participants in the protocol aware of their position in the chain or any detailsabout the other neighbourhoods, except for the immediate next one to which they are push-syncing their chunk.
+Neither are the participants in the protocol aware of their position in the chain or any details about the other neighbourhoods, except for the immediate next one to which they are push-syncing their chunk.
 
@@
 
-A powerful adversary could potentially infiltrate every neighborhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it).
+A powerful adversary could potentially infiltrate every neighbourhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it).
 They could also keep logs of what has been uploaded in what order, which would allow it to serve up deleted content on demand.

### 2025-09-11 for-submission.tex
revision submitted to acm  tweb special issue

%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,manuscript')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-manuscript.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%  
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[manuscript,screen,review]{acmart}
%%
\usepackage[hang,flushmargin]{footmisc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{scalerel}
\newcommand\defeq{\overset{\textrm{\tiny  def}}{=}}
\newcommand\Xor{\veebar}
\newcommand\xor{\text{\small$\veebar$}}
\DeclareMathOperator*{\XOR}{\scalerel*{\Xor}{\text{\small$\sum$}}}
\newcommand{\node}{\nu}
\newcommand{\Chi}{\text{\large$\chi$}}
\newcommand{\chimin}{\overline{\text{\large$\chi$}}}
\newcommand{\idx}[1]{\texttt{[}\/#1\/\texttt{]}}
\setlist[itemize]{label={---}}
% the following code puts a line break after a description label even if label is shorter than leftmargin
\renewcommand*{\descriptionlabel}[1]{\hspace\labelsep\parbox{\linewidth}{#1}}
\newlist{labelledlist}{description}{2}
\setlist[labelledlist]{style=nextline, leftmargin=2em, before=\renewcommand*{\descriptionlabel}[1]{\hspace\labelsep}}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%\setcopyright{acmlicensed}
%\copyrightyear{2018}
%\acmYear{2018}
%\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
%\acmConference[Conference acronym 'XX]{Make sure to enter the correct
%  conference title from your rights confirmation email}{June 03--05,
%  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
%\acmISBN{978-1-4503-XXXX-X/2018/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{A dream come true: deletable content in immutable storage}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Viktor Tr\'{o}n}
\authornote{Corresponding author}
\email{viktor@ethswarm.org}
\orcid{0009-0003-0222-6855}
% \affiliation{%
%   \institution{Swarm Research Division}
%   \city{Neuch\^{a}tel}
%   \country{Switzerland}
% }

% \author{D\'{a}niel A. Nagy}
% \email{daniel@ethswarm.org}
\affiliation{%
  \institution{Swarm Research Division}
  \city{Neuch\^{a}tel}
  \country{Switzerland}
}
\author{Henning Diedrich}
\email{hd@lexon.org}
\orcid{0009-0008-0798-0132}
\affiliation{%
    \institution{Ethersphere Core Task Force, KSCC}
    \city{Berlin}
    \country{Germany}
}
%%
%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Tr\'{o}n \& Diedrich}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
This work introduces DREAM, a novel construct that enables deletable content in otherwise immutable, decentralised storage systems such as Swarm. As an equitable and feasible response to the challenge of data destruction, DREAM reconciles the inherent persistence of content-addressed storage with the practical needs, user preferences, and legal requirements of permanent access revocation.
The paper contributes a new deletion model based on access revocation, and presents a mechanism that allows data-sharing users to grant—and later rescind—access to specific data for specific consumers. The solution leverages Swarm’s distributed storage architecture to introduce a procedural way to retrieve decryption keys, realised by a network protocol. The resulting system achieves the desired properties of deniability, revocability, expirability, addressability, and malleability without requiring trusted intermediaries or complex cryptographic primitives. The approach preserves the censorship resistance of decentralised systems allowing for sovereign control over data access by the original uploader, but not by any other party, thus avoiding any re-centralisation in order to achieve deletion. Security analysis shows that access revocation remains effective even under pervasive adversarial control.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
% \begin{CCSXML}
% <ccs2012>
%  <concept>
%   <concept_id>00000000.0000000.0000000</concept_id>
%   <concept_desc>Networks, Network algorithms</concept_desc>
%   <concept_significance>500</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>00000000.00000000.00000000</concept_id>
%   <concept_desc>Networks, Network protocols</concept_desc>
%   <concept_significance>300</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>00000000.00000000.00000000</concept_id>
%   <concept_desc>Information systems, Information storage systems</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>00000000.00000000.00000000</concept_id>
%   <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
% </ccs2012>
% \end{CCSXML}

% \ccsdesc[500]{Networks~Network algorithms}
\ccsdesc[500]{Networks~Network protocols}
\ccsdesc[500]{Information systems~Information storage systems}
\ccsdesc[500]{Information systems~World Wide Web}
\ccsdesc[500]{Security and privacy~Database and storage security}
% \ccsdesc[500]{Social and professional topics~Technology policy}
%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Swarm, Web3, decentralisation, data deletion, immutability}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

This paper is structured as follows:
Section~\ref{sec:intro} acknowledges the need for data deletion and explains why, in practice, deletion is  typically implemented and understood as removal of access. It presents the challenges of realising deletion in decentralised storage systems, exploring the tension between censorship resistance and access control.
Section~\ref{sec:deletion} formalises the concept of deletion in the context of content-addressed networks like Swarm and derives the criteria required for meaningful access revocation.
Section~\ref{sec:construction} describes the technical construction of the DREAM protocol, detailing its architecture, design rationale, and its five core properties.
Section~\ref{sec:analysis} analyses the protocol’s correctness and behaviour under realistic network assumptions.
Section~\ref{sec:security} provides a security model and evaluates the system’s resilience against adversarial behaviour, in particular, under conditions of pervasive network compromise.
Section~\ref{sec:constraints} discusses the prerequisites for implementing DREAM and argues that adopting it would require only modest effort.
Section~\ref{sec:conclusion} concludes by summarising the findings and highlighting the broader implications of user-friendly, access-controlled deletion in the context of immutable storage.

\section{Introduction} \label{sec:intro}

The continuously increasing digitisation of data and the widespread use of cloud services in recent years have raised well-justified concerns regarding privacy and control of personal information \cite{voigt2017gdpr}. 
In  response, regulators often impose requirements for data removability  with little regard for technical or practical feasibility. Beyond the truism that it is virtually impossible to make material once seen unseen, legal and economic realities often mean that ostensibly `deleted' data is merely hidden rather than physically erased. Social media platforms, for example, list a plethora of exceptions to reserve the right to not actually delete the data of a `deleted' account \cite{facebook2025privacy}. And while they are generally obliged by law to erase the data, some of their valid counter-arguments are themselves regulatory and legal requirements. Even at the technical level, a local, high-performance database like Facebook's workhorse, Apache Cassandra, does not actually delete data \cite{apache2025cassandra} when it is flagged as `deleted,' purely for the mundane reason of higher performance, i.e. lower cost.

But because entities that operate centralised infrastructure for digital publishing typically have full control over the systems that serve their content, they can carry out the effective `deletion' of data by simply \emph{denying access} to it.

No wonder then that centralised gatekeeping is appealing from a regulatory standpoint: it can be summoned as a convenient instrument of law enforcement. And sure enough, in case legal action is initiated, the law routinely forbids the destruction of data that could constitute evidence. The centralized gatekeeper is thus morphed into a responsible party with an effective, if paradoxical, course of action: not serving censored content, but preserving it.

Even though this can give the unwary a false sense of privacy, it illustrates a deeper issue with censorship: publishing platforms possess the technological means to filter content, and centralised hosting makes it easy and economical to exert control over it. What may have started out as a benign and sensible measure of content curation can gradually transform into extant censorship \cite{chaum2021informationage,gillespie2018custodians}. This is all the more problematic with social platforms \cite{barnett2019platforms,klonick2018newgovernors} that have gained quasi-monopolistic status as a result of network effects \cite{zuckerman2014newcensorship}. This allowed governments to co-opt moderation tools in service of political agendas \cite{facebook2023pressure}. And due to the high costs associated with switching platforms, content creators increasingly find themselves 'deplatformed' \cite{zhang2021deplatforming} for violating the arbitrary and frequently changing publishing rules of the platforms they rely on. The ability to identify (central) hosts and coerce them through legal means to deny access to content provides authorities with a cheap and easy tool to curtail freedom of speech.

Recognising how easily technically motivated solutions can be put in service of propaganda, manipulation and surveillance has motivated the work of hackers for decades. The vision to enable censorship-resistant publishing has been at the forefront of the decentralisation efforts they initiated. Since  the early days of the \emph{cypherpunk movement} in the 1980s, \emph{censorship resistance} has become a foundational principle of peer-to-peer storage networks, such as Swarm \cite{swarm2021whitepaper}.

In the decentralised paradigm of Web3, there is, as per its founding mission, no longer a single operating entity controlling publishing platforms or hosting infrastructure. On the one hand, this renders censorship%
%
\footnote{Censorship is understood as the act of a third party to make data unavailable. In practice, this is the case when said 'single operating entity controlling publishing platforms or hosting infrastructure' makes content inaccessible. Censorship is not the deletion of something previously posted by its original poster or uploader, which, when enabled in whatever form, does not therefore violate censorship-resistance.} 
%
infeasibly costly, as intended. Content posted in a Web3 environment \textit{can} disappear if it is not kept up. But the data is censorship-proof in that its disappearance cannot be hastened by anyone other than the original poster. With a decentralized, immutable platform, even the poster cannot make data disappear at will though. And thus, the potentially unintended yet permanent exposure of personal data presents a serious concern for many individuals. Unless one accepts this as the  price to be paid for censorship resistence, there is a need for solutions that can restore the sense of security that centralised gatekeepers provide \cite{geambasu2009vanish}.

Prima facie, the intent to introduce a mechanism to delete content from \emph{immutable} platforms seems futile. However, immutability does not imply that content cannot be updated, nor that the network itself cannot change; it simply means that once data is stored, it cannot be \emph{actively} deleted.

To overcome this apparent paradox, we need to acknowledge that 1) the requirement of  `deletion' normally does not refer to physical destruction of offending content but to the removal of access to it; combine this with the fact that 2) data can be stored encrypted, making loss of access equivalent to `losing the key'; that 3) a deterministic random generator can be provided trustlessly  to create a key through a social process defined by a network protocol, as explained below; that 4) immutable storage environments allow for adding data; and 5) that the act of adding seemingly unrelated content can change the protocol's output, altering the key derived from the same seed.

Armed with these ingredients, we can arrive at a mechanism to \emph{effectively} delete data from an immutable, distributed network, and comply with both user sentiment and legal requirements; and with this, neutralise a perceived advantage of traditional, centralised data storage designs.

\section{Deletion and revoking access}
\label{sec:deletion}

Requirements of removal of information in the sense of erasing it from all physical storage devices are both unenforceable and impractical \cite{ohm2010broken}. Even the most rigorous data protection audits do not require the erasure of offending data from backup tapes. In general, the tacit assumption is that information ordered to be removed should become \emph{inaccessible} through \emph{typical, original, and precedented methods of access}.

In what follows, we formulate what we think is the strongest meaningful definition of deletion applicable to permissionless, immutable, decentralised storage systems and offer a construction that implements it.  Importantly, this approach is purely technical: it relies on the capabilities and costs of primary actors, rather than on procedural measures that impose obligations on intermediaries to respect those actors' rights. In other words, it operates completely algorithmically in a trustless, distributed network, without expecting anyone to comply either out of altruistic motive or under punitive threat.

The primary actors in this context are users who wish to share content (the \emph{uploaders}) by granting read access to a number of data consumers (\emph{downloaders}).  Granting read access is defined as providing a canonical \emph{reference} to the content, allowing the system to retrieve the complete information intended for disclosure. Revoking access means that this canonical reference stops working.

It is obvious that any party that is privileged to access information could store, re-code, and potentially disseminate it, so that the content can be made accessible at a later point in time, which practically bypasses any process that would qualify as deletion (or removal of access). As there is no guaranteed protection against such adversity, any legally and socially useful notion of deletion invariably defines a narrower case: taking away \emph{the viability to replay the same access method} at a lower cost than at least the full cost of storing the content.%
%
\footnote{That is, the total storage cost paid for the full size of the content starting from the time that access was revoked up until the attempted breach.}
%
At most, this could mean the full cost of storing all content and/or all change logs of the publishing system holding the specific content, since it is impossible to know in advance which piece of data may become relevant in the future.  For example, deletion on a social media platform only means that the original link no longer works; even if others may have downloaded and reposted the content elsewhere under a new link.%

Consequently, we will explore deletion as a scheme for uploading content with access revocation that meets the following criteria:

\begin{labelledlist}[noitemsep]
    \item[\emph{specialisation}] The uploader is able to choose at the time of publishing a specialised construct that will enable access revocation later. From a user's perspective, content that is meant to be reliably deleted should be marked or constructed as such at the time of upload. The cost of uploading and preserving such content may be higher than the cost of regular, censorship-resistant but non-deletable content.

    \item[\emph{sovereignty}] Deletables presuppose access control, i.e. they are only available to a specific set of recipients. The uploader is the unconditional owner of the deletable data and holds the exclusive means to selectively revoke access from any party that was previously granted access. In other words, the uploader's credentials are necessary to delete their own deletable content, and no one else can delete it.
%
    \item[\emph{security}] After access is revoked, a grantee is unable to access the content using the same reference or any other cue shorter than the deleted content itself. That is, unless the downloader has stored at least as much  as the size of the deleted content itself, they will have no way of retrieving it.
%
\end{labelledlist}

Note that in the Swarm network, uploaded content may eventually be \emph{forgotten}: if nobody pays for its storage and the content is not frequently accessed, it will be garbage-collected. However, content no longer paid for to be stored
%
\footnote{In the context of Swarm, chunks with \textit{expired postage stamps}.}
%
cannot be regarded as reliably deleted: the requirements of sovereignty and security are not strictly fulfilled.

\section{Construction}
\label{sec:construction}

The goal of this section is to arrive at a formal construction of a revocable access model for the distributed storage network Swarm. We will base our exploration on \emph{chunks}, the fundamental fixed-size storage units of Swarm's distributed storage model, DISC \cite[\S 2]{tron2025BOS}. The proposed \emph{DREAM} construct implements a deletable content storage and access model that fulfills the requirements of specialisation, sovereignty, and security.

\subsection{Synopsis}

The general approach of the contribution is that access to content can be revoked by taking away the decryption key. For this to be possible, the key itself must be of volatile nature. In the proposed protocol, such a key ($k$) is constructed for the grantee by data flowing  through the network of Swarm nodes  (see figure~\ref{fig:dream-path}). Starting out as a chunk-length piece of data generated from a random seed ($g$) it is sent to its destination wrapped as a content-addressed chunk beginning its journey of transformation following a \emph{dream path}. At each step, the chunk is picked up by a storer node in the target neighbourhood, the content is modulated using local chunk data in the node's storage as input ($\Delta$ function) then once-again wrapped as a chunk, its content address determining in which neighbourhood the subsequent step will be performed. After an arbitrary but fixed number of steps ($n$), the dream path leads back to the grantee and the key $k$ is extracted and can be used to reconstruct the deletable content $C$ by decrypting its encrypted form $C'$ that is stored in the network with $k$.

The uploader of deletable content does not need to play this script of collective construction: critically, because each step takes input from data previously uploaded using the same \emph{postage stamp} owned by the uploader, they have all the data needed to simulate the protocol's calculations. Once the owner mines a key by finding a path, the deletable content ($C$) is encrypted, and the resulting encrypted content ($C'$) is then uploaded. To share a link to the content, the uploader shares a dream reference, a tuple $\langle r,b,g\rangle $ consisting of the Swarm reference $r$ to the encrypted content $C'$, a postage stamp batch reference to the batch ID $b$, and the generator seed $g$. The grantee retrieves the encrypted content by the Swarm reference $r$. The batch id $b$ and the generator seed $g$ enable the grantee to obtain the key from the network, triggering the path traversal to re-create the dream key $k$. The content plaintext $C$ can be obtained by decrypting $C'$ with the key $k$.  Since the key itself has the same length as the content, there is no incentive to store $\langle r,k\rangle $ as an alternative, because it would take more storage than simply storing the content $C$.

Importantly, while the key to the data can be accessed through the network, it is never stored in it. This is what makes revocation possible. Since the owner of the stamp (the uploader) controls that dataset, by virtue of uploading extra data to any node's local storage, they can change the spice coming from local data (figure \ref{fig:deflected}). This is sufficient to deflect the path which renders the key inaccessible, and thus effectively delete their content from the network.

\subsection{Dream: deniable, revocable, expirable, addressable, malleable}

The use of the word `dream' alludes to the counterintuitive nature of the finding that deletion is even possible in an \emph{immutable} context.  As a mnemonic acronym, it resolves to the 5 \emph{dream attributes} that the proposed construct displays:

\begin{labelledlist}

    \item[\textbf{D} \emph{deniable}] The dream key constructed by the protocol is used for both encryption and decryption. Since any content chunk (in fact, any arbitrary content) could be encrypted using it, the key's association to any specific content is plausibly deniable (see section \ref{sec:security}).

    \item[\textbf{R} \emph{revocable}] Access granted through dream keys is revocable. Revoking access from all parties, including oneself, is considered deletion. What is actually revoked is the ability to retrieve the key, which invalidates the reference $\langle r,b,g\rangle $ to the decrypted content.

    \item[\textbf{E} \emph{expirable}] The scheme allows for one-time use in this design. By strategically limiting the Swarm postage stamp supply associated with the key for its (re-)creation, the \emph{key} can be made to only be retrievable once.

    \item[\textbf{A} \emph{addressable}] Access can be granted to nodes in any \emph{neighbourhood} $h$ by mining a dream path of length $n$ such that the last chunk falls in $h$,%
%
    \footnote{Swarm neighbourhoods are groups of nodes which are responsible for sharing the same chunks \cite{swarm2021whitepaper}.}
%
    where only clients operating a node within a particular overlay address range are able to access the content.

    \item[\textbf{M} \emph{malleable}]   The construct is resilient to churn and dynamic changes in network size; it is reusable across independent grantees and upgradeable.

\end{labelledlist}

The dream protocol can be built on top of Swarm's DISC%
%
\footnote{DISC (Distributed Immutable Storage of Chunks) is a storage solution developed by Swarm based on a modified interpretation of a Kademlia DHT (Distributed Hash Table) which has been specialized for data storage. Swarm's implementation of a DHT differs in that it stores the content in the DHT directly, rather than just storing a list of seeders who are able to serve the content. This approach allows for significantly faster and more efficient retrieval of data.}
%
APIs as a pure second-layer solution. Despite its rich feature set, the scheme does not use complex cryptographic primitives, but instead leverages the interplay of various component subsystems.

\subsection{Chunk upload in Swarm}

When content is uploaded to Swarm, as per its standard operation, the local Swarm client splits the data into 4-kilobyte fragments, wraps each of them with some metadata and an address into a \emph{chunk}, and pushes them to the network using the \emph{push-sync} protocol. This protocol governs the transmission of newly entered chunks to the neighbourhood where they are to be stored.

A Swarm chunk $c$ is an association $\langle a, c'\rangle$ of a 32-byte address $a$ and chunk content $c'$. While Swarm has other chunk types, the DREAM protocol focuses on the \emph{content-addressed chunk} (CAC). CACs assume that their content $c'$ is the chunk data $p$ of length limited to 4096 bytes prepended with associated metadata $m$. Content-addressed chunks attain their integrity by having their address deterministically derived from their content (including metadata and chunk data) with the help of a one-way uniform digest, a \emph{hash function} $\mathit{H}$.%
%
\footnote{The BMT hash is the Keccak256 hash of  the metadata prepended to the root hash of the \emph{binary Merkle tree} with a Keccak256 base hash over 32-byte segments of the chunk data. This particular choice of $H$ does not affect the correctness of our construction.}
%
%

\begin{eqnarray}
\mathit{Data} & \defeq & \mathit{byte}\{,4096\}\\
\mathit{Meta} & \defeq & \mathit{byte}\{8\}\\
\mathit{Address} & \defeq & \mathit{byte}\{32\}\\
\mathit{Content} & \defeq & \mathit{Data} \times \mathit{Meta}\\
\mathit{Chunk}&\defeq& \mathit{Address}\times \mathit{Content}\\
% \mathit{Batch}&\defeq& \\ 
\mathit{CAC} &:& Data \times Meta\to  \mathit{Chunk}\\
\mathit{CAC}(p,m)&\defeq&\langle \mathit{Keccak}(m\oplus\mathit{BMT-root}(p)),m \oplus p\rangle
\end{eqnarray}

We also define:
% %
\begin{eqnarray}
\mathit{Address} &:& \mathit{Chunk} \to \mathit{Address} \\
\mathit{Address}(\langle a, c'\rangle) &\defeq& a\\
\mathit{Data} &:& \mathit{Chunk} \to \mathit{Data} \\
\mathit{Data}(\langle a, c'\rangle) &\defeq& c'\idx{8:}
\end{eqnarray}
% %

The delivery of a new chunk to the node where it is to be stored can be seen as a request, to which one of the nodes closest to the chunk's hash position in the DHT eventually responds with an acknowledgment of custody.
Swarm's network protocols use \emph{forwarding Kademlia}, meaning that requests are routed to their destination through a series of hops successively relayed by the \emph{forwarding} nodes with ever increasing proximity to the target address, i.e. the chunk's address $a$. 
The response is passed back via the same route as the original request (\emph{backwarding}) keeping the identity of the request originator private, see figure~\ref{fig:pushsync}.

\emph{Proximity} is the inverse of the xor distance metric \cite{maymounkov2002kademlia}, while 
\emph{proximity order} $\mathit{PO}(x,y)$ returns the integer part of the logarithm of the xor distance, which can be calculated as the number of matching initial bits of the two addresses $x$ and $y$.%
%
\footnote{As an example, the proximity order $\mathit{PO}$ of 00101100 and 00110011 is 3, because the two 8-bit integers share the prefix 001, which is three bits long. }
%
Both proximity and proximity order apply to the address space shared by chunks and peers and play a crucial role in Swarm's peer-to-peer routing.
Conversely, an address range is specified by a bit sequence, which defines a \emph{neighbourhood} of chunks (reserve), or peers. Given an address $a$ and a non-negative   integer depth $d$, the shared \emph{prefix} of all addresses in the range can be specified as the initial $d$ bits of $a$, called a neighbourhood of \emph{depth $d$ designated by $a$}. In practice, `neighbourhood' refers to a cluster of nodes that are incentivised to relay as well as replicate and retain all chunks whose addresses share such prefix with the addresses of nodes in the cluster.%
%
\footnote{Due to the Kademlia connectivity, the route to the node whose address is closest to the chunk can be found using local decisions by the nodes along the path. Also, the forwarding-backwarding routing scheme makes it possible to do bandwidth accounting on a peer-to-peer basis and facilitates micropayment settlements on a repeated-dealings basis with a subset of quasi-permanent peers that is logarithmic in the size of the network.}

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{figs/push-sync.pdf}
  \caption{Push-syncing. The protocol is responsible for transferring newly uploaded chunks to the neighbourhood of peers where they belong based on address proximity to the chunks' content addresses. Uploader $U$ posts a \emph{chunk} to a local peer connection  $F_0$, which forwards it to a node closer to storer $S$, etc. The protocol's request—response scheme uses the forwarding/backwarding Kademlia relay.
}
  \Description[Push-Syncing]{The protocol that takes care of pushing a chunk to its destination based on its address.}
\label{fig:pushsync}
\end{figure}

Eventually, every chunk uploaded to Swarm is push-synced and forwarded by relaying nodes toward its address, where it is stored in the nodes' \emph{reserve}. The length of the shared prefix prescribed to accept a neighbourhood as target area is called storage depth $\mathit{depth}_t$. It depends on the amount of storage space nodes must dedicate for their reserve and the overall volume uploaded to the network. The depth is incremented as the volume doubles and decremented as the volume halves.

This procedure requires spam protection to prevent mass injection of random chunks into the network, which could expunge valuable chunks by triggering garbage collection on nodes that are pushed into capacity shortage. By requiring \emph{postage stamps}, Swarm imposes an upfront cost to uploads as an effective measure against such spamming. A chunk is accepted into a node's reserve only if it has a valid postage stamp attached. Revenue resulting from postage stamp purchases constitute the reward pot for the redistribution game, compensating storage nodes for data retention.

\subsection{High-level description}

The central construct of this contribution is called the \emph{dream key} ($k$). It is a random byte sequence the size of a chunk's data ($p$), and it acts as a symmetric key (for encryption and decryption) for the deletable content $C$. Revoking access to the dream key $k$ realises the deletion of the content $C$ that was encrypted using it. The critical question we are addressing is how access could be revoked in an immutable storage network. The solution is to never store the dream key directly, but to have it collectively created by a set of nodes on demand, according to the protocol described below.\footnote{As mentioned, since the dream key is as large as the data it encrypts, there is no advantage for anyone to store it instead of the decrypted data itself.}

The creation of the key starts with a chunk-sized (4096-byte) sequence $p_0$ deterministically generated from a key-sized (32-byte) seed $g$ by a pseudo-random generator function $\mathcal{G}\idx{4K}$.%
%
\footnote{For example, the Keccak sponge function used throughout Ethereum for hashing does have this capability. Alternatively, the block cipher encryption using the seed as initial nonce can be applied to a fixed constant chunk, such as all zeros.}
%
The protocol then sends this first iteration stage of the key as a content-addressed chunk (CAC) $c_0 = \langle a_0, m \oplus p_0\rangle$ to the network, $m$ being metadata that helps with the key generation. As per Swarm's push-sync protocol, a node $\node_0$ in the neighbourhood designated by the chunk address $a_0$ is guaranteed to eventually receive this chunk. Now, the dream protocol prescribes that the closest node must update the key $p_0$ to $p_1$ in a specific way, using $b$ and the \emph{key update function} $\Delta$; and then, wrapped as a new chunk $c_1 =\langle a_1, m \oplus p_1\rangle$, sent it to the network again towards the neighbourhood designated by its new content address $a_1$. The next node that updates the dream chunk is considered the next node on the dream path ($\node_0\ldots\node_n$ in figure~\ref{fig:dream-path}).%
%
\footnote{Note that, due to forwarding Kademlia, routing is realised by relaying the messages via multiple hops using keep-alive peer connections as the transmission channel. However, for the dream protocol, these intermediate forwarding nodes are not relevant.}
%
%
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/dream-path-2.2.pdf}
    \caption{Left: The path of the solid arrows represents the dream path, a series of neighbourhoods the dream protocol needs to traverse. The grantee posts the chunk data $p_0$ generated from $g$ wrapped as chunk $c_0=\langle a_0, m \oplus p_0\rangle$ no different from any other upload to Swarm. Through multiple forwarding hops, $c_0$ arrives where it would be stored according to $a_0 = \mathit{Address}(c_0)$. The dream key is created as the chunk-long piece of data journeys through the neighbourhoods designated by $\mathit{Address}(c_0),\ldots,\mathit{Address}(c_n)$, picked up by the closest nodes $\node_0 \ldots \node_n$. Eventually, at step $n+1$, $c_{n}$ is received back by grantee $G$ at address $a$ who can extract the dream key $k=\mathit{Data}(c_n)$.
    Right: For each step $0\le i<n$ on the way node $\node_i$ updates $c_i$ using the update function $\Delta$. Information locally available to nodes specific to that neighbourhood ($\mathit{BB}_t(b,c_i)$) serves as input to the key update function ($\Delta$). 
    The ($\emph{1.}$) incoming chunk $c_{i}$ is used to ($\emph{2.}$) find in the reserve the set $\mathit{BB}_t(b,c_i)$, i.e. those chunks that have a valid postage stamp issued under $b$ and fall into the batch bucket designated by the input chunk address $\mathit{Address}(c_i)$. These chunks' data are then XOR-ed together with the input chunk data $\mathit{Data}(c_{i})$ to create $p_{i+1}$, which is then wrapped into the chunk $c_{i+1}$ and uploaded ($\emph{3.}$).}
    \Description[Dream Path]{The path that is how the network stores a dream key.}
\label{fig:dream-path}
\end{figure}

Every node that receives the evolving dream chunk wraps it using data from its \emph{reserve} at the time ($BB_t$) as input for $\Delta$, i.e. from the chunks that it is paid to store. If the protocol is initiated more than once, conditions may be different due to network and reserve changes. This sensitivity of conditions to time is indicated with a subscript $t$. In fact, the key invention is to use the ability of the uploader to change the reserves of a neighbourhood, to thereby derail a dream path that previously resolved.

But as long as the initial key $p_0$ and the metadata $m$ are the same {and} the relevant parts of the reserve $BB$ remain unchanged, the system ends up generating the same dream key, which allows the grantee to decrypt and read the deletable content.

Critically, by leaving control over these conditions in the hands of the uploader, they are the only party that can actively block the grantee's ability to calculate the key.
%
%
\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.85\textwidth]{figs/dream-path-2.2-deflected.pdf}
  \caption{Deletion. The dream path is shown as `deflected' by a new chunk $u$ stamped by postage batch $b$ sent by uploader $U$ ($\emph{1.}$).
  An uploader may alter the dream path by inserting a chunk at time $t'$ into the relevant batch bucket in any step $0\le i< n$. As a result, $\mathit{BB}_{t}(b,c_i)\neq \mathit{BB}_{t''}(b,c_i)$ for all $t<t'<t''$ for at least those steps $i$ effected. This change will result in a different output chunk $c_{i+1}$ at step $i$, whereby the dream path is \emph{deflected} and will not terminate with the grantee who is now unable to retrieve the dream key to access the `deleted'  content.}
  \Description[Update Function]{The function that describes how each node on the dream path incrementally changes the incoming dream key.}
\label{fig:deflected}
\end{figure}
%
%
Overall, this makes retrieving a deletable chunk a process involving multiple nodes. Because calculating one part of the reference to a deletable chunk requires sorting through a set of immutable chunks controlled by the uploader, the uploader's ability to change this underlying set—by adding to it—provides the mutability to take access away (figure~\ref{fig:deflected}).

\subsection{Making up the dream}

The specific way that each node in the dream path wraps the chunk that will become the dream key before sending it on is described in the following (see figure \ref{fig:dream-path}). Let the 32-byte hash $b$ be the ID of a postage batch, i.e. a set of stamps, owned by uploader $U$. The node receiving a dream chunk looks $b$ up in their batchstore by the prefix extracted from the chunk metadata. We assume that this prefix will always match $b$'s ID uniquely, as long as $b$ is valid at the time $t$ of receiving the chunk. Chunks in a bucket share a prefix corresponding to the \emph{uniformity depth} $\upsilon(b)$. In order to ensure that nodes can always detect doubly signed storage slots, Swarm stipulates that neighbourhood depth must not exceed uniformity depth for the entire period that the stamp is valid:
%
\begin{eqnarray}\label{eq:bb}
\forall b\in\mathit{Batch}, \forall t,  b \text{ is valid at }t\rightarrow\upsilon(b)\ge \mathit{depth}_t
\end{eqnarray}
%
Let $\mathit{BB}_t(b,a)$ represent the set of chunks belonging to batch $b$ at time $t$ and fall into the bucket  designated by $a$.
Since the chunks belonging to a bucket of a batch all share a prefix longer than the prefix shared by the storer nodes in the neighbourhood, it is expected that all chunks in a batch bucket designated by $a$ are stored in the reserves of nodes in the relevant neighbourhood designated by $a$.
It follows from equation \ref{eq:bb} that neighbourhoods' reserves must contain entire buckets:
%
\begin{eqnarray}
\forall a\in\mathit{Address}, b\in\mathit{Batch},  \forall t,  b \text{ is valid at }t\rightarrow
\mathit{BB}_t(b,a)\subseteq\{c\mid\mathit{PO}(\node,c)\ge\mathit{depth}_t\}
\end{eqnarray}
%
The \emph{key update function} $\Delta$ updates the key by bitwise XOR-ing it with each chunk data belonging to batch $b$ at block height $t$ and falling into the bucket designated by the input chunk's address.
\begin{eqnarray}
{\Delta}&:&
\mathit{Batch}\times\mathit{Chunk}\to\mathit{Data}\\
{\Delta}(b,\langle a_i,c_i\rangle)&:&\mathit{Data}(c_i)\xor\XOR_{c\in\mathit{BB}_t(b,a_i)}\mathit{Data}(c)
% {\Delta}(b,\langle a_i,c_i\rangle)&:&\mathit{Data}(c_i)\xor\XOR_{c}^{\mathit{BB}_t(b,a_i)}\mathit{Data}(c)
% {\Delta}(b,c_i)&:&\mathit{Data}(c_i)\xor\XOR_{c\in \mathit{BB}_t(b,\mathit{Address}(c_i))}\mathit{Data}(c)
\end{eqnarray}

If the bucket of batch $b$ designated by $a_i$ is empty, then $\Delta$ returns $\mathit{Data}(c_i)$.
In any case, the update function is locally computable by any storer node within the neighbourhood designated by $a_i$, as it will have all required input (reserve chunks in $\mathit{BB}_t(b,a_i)$) stored.

Applying the update function $\Delta$ to its own output, by the next node, we can define the \emph{dream path} $\Pi$ as follows:
%
\begin{eqnarray}
\Pi&:&  \mathit{Batch}\times\mathit{Address}\to\mathit{Chunk}\{n+1\}\\
\Pi(b,g)&\defeq&c_0,\ldots, c_n\text{ such that }\\
c_i &\defeq&\mathit{CAC}(p,b\idx{0:8})\text{ where }
p=\begin{cases}
\mathcal{G}\idx{4K}(g)&\text{if }i=0\\   
{\Delta}(b,c_{i-1})&\text{otherwise}             
\end{cases}
\end{eqnarray}
%
Finally, a \emph{dream pair} $\langle g,p_n\rangle$ is a particular pairing of a seed $g$ that is the input to the generator function creating the initial key $p_0$; and the dream key $p_n$, the result of $n$ applications of the key update function.

The starting point, seed $g$, such that the dream path ends after $n$ steps in a neighbourhood $a$ of nodes that $U$ desires to be the potential exclusive receivers of the deletable chunk's payload, cannot be calculated directly. It is instead worked out by trial-and-error, or in other words, $g$ is mined.

Therefore, given a dream path of length $n$, a grantee overlay address $a$, and a batch $b$, the uploader needs to find the generator $g$ such that the last chunk of the dream path falls within $a$'s neighbourhood.
%
\begin{eqnarray}
\mathit{dream}&:&\mathit{Address}\times\mathit{Batch}\times\mathbb{Z}^{+}\to\mathcal{P}(\mathit{Address}\times\mathit{Data})\\
\mathit{dream}(a,b,n) &\subset &\mathit{Address}\times\mathit{Data}\\
\langle g, k\rangle \in \mathit{dream}(a,b,n) 
&\Leftrightarrow
& k=\mathit{Data}(c_n)\land c_n=\Pi(b,g)\idx{n+1}\land\mathit{PO}(a,\mathit{Address}(c_n))\geq d
\end{eqnarray}
%
Since the hash function $\mathit{H}$ used in calculating $a_n=\mathit{Address}(c_n)$ is uniform, the chance of $\mathit{PO_{256}}(a,a_n)\geq d$ is 1 in $2^{d}$. As a consequence, it is feasible for the uploader to find $g$ for any sufficiently low depth $d$,  which, in turn, allows them to calculate the dream key $k$ for a dream path ending at $a$.%

\subsection{The Dream Protocol}

To calculate the dream key, grantees must rely on the network, where each recursive step $i$ of the calculation for all $0\le i\le n$ must be performed by the node closest to the input chunk's address $a_{i}$.

In order to guarantee the correct termination, the following criteria must be fulfilled:
\begin{itemize}[noitemsep]
\item[---] all buckets of batch $b$ designated by $a_i$ for all $0\le i< n$ must be non-empty,
\item[---] the length of the prefix shared by the addresses of chunks in the same bucket must be greater than the storage depth $d$ of the network,
\item[---] the chunk $c_{i}$ of each step $0\le i\le n$ must be sent to the neighbourhood designated by  $a_{i}=\mathit{Address}(c_{i})$,
\item[---] nodes at each step $0\le i<n$ of the dream path must both compute the key update and push the output key as a chunk to the network.
\end{itemize}

% \begin{figure}[htbp]
%   \centering
%     \includegraphics[width=0.6\textwidth]{figs/dream-protocol-1.16.pdf}
%   \caption{The dream protocol. Assuming uploader $U$ mined a \textit{dream pair} $\langle g,k\rangle$ of length $n$ for grantee at overlay $a$ and nodes in the network listen to \emph{dream chunks}, content-addressed chunks whose metadata $\beta$ is a prefix of a valid batch ID $b$. Once they receive dream chunk $c_{i}$, and also look up $b$ using prefix $\beta$, they then find the chunks that are stamped with $b$, fall into the bucket  of the input chunk address $a_{i}$. The updated key output $k_{i+1}=\Delta(b,c_i)$ is then wrapped together with $\beta$ as a content-addressed chunk $c_{i+1}=\mathit{CAC}(k_{i+1},\beta)$ and uploaded to the network to end up in the neighbourhood designated by $\mathit{Address}(c_i)$ as the next step on the dream path. The node that is closest to $\mathit{Address}(c_i)$  then calculates the next step based on this input. If the protocol is followed up to $n$ steps, then the target node at address $a$ receives $c_{n}$ from which it can extract $k=\mathit{Data}(c_{n})$.}
%   \Description[Dream Protocol]{The protocol requires a series of incentivised, local computation and must produce a chunk-sized encryption key.}
% \label{fig:protocol}
% \end{figure}

We define the dream network protocol as follows: Assuming uploader $U$ mined a \textit{dream pair} $\langle g,k\rangle$ of length $n$ for grantee at overlay $a$ and nodes in the network listen to \emph{dream chunks}, content-addressed chunks whose metadata $\beta$ is a prefix of a valid batch ID $b$. Once they receive dream chunk $c_{i}$, and also look up $b$ using prefix $\beta$, they then find the chunks that are stamped with $b$, fall into the bucket  of the input chunk address $a_{i}$. The updated key output $k_{i+1}=\Delta(b,c_i)$ is then wrapped together with $\beta$ as a content-addressed chunk $c_{i+1}=\mathit{CAC}(k_{i+1},\beta)$ and uploaded to the network to end up in the neighbourhood designated by $a_{i+1}=\mathit{Address}(c_{i+1})$ as the next step on the dream path. The node that is closest to $a_{i+1}$  then calculates the next step based on this input. If the protocol is followed up to $n$ steps, then the target node at address $a$ receives $c_{n}$ from which it can extract $k=\mathit{Data}(c_{n})$.

The deletable dream chunk is constructed as follows: if uploader $U$ wishes to grant downloader $D$ (at overlay address $a$) revocable access to  chunk content $C$, then $U$ chooses a postage batch $b$ that it owns, and that is \emph{not completely filled}.
$U$ then by trying random seeds $g$, mines a dream pair $\langle g, k\rangle\in\mathit{dream}(a,b,n)$,
calculates the ciphertext $C'=C \xor k$ and uploads it to Swarm, obtaining the parity reference $r=\mathit{H}(C')$. Now $U$ creates the \emph{dream chunk reference} $\mathit{ref}(C)=\langle r,b ,g\rangle $, which must be privately shared with grantee $D$.

\section{Analysis}
\label{sec:analysis}

% \subsection{Retrieval}
% \label{def:retrieval}
Downloader $D$, in possession of a dream chunk reference $\mathit{ref}(C)=\langle r,b,g\rangle $ calculated by $U$, constructs $p_0=\mathcal{G}\idx{4K}(g)$, wraps it as the initial chunk $c_0=\mathit{CAC}(p_0,b\idx{0:8})$, and uploads it to Swarm. If and when $D$ receives the dream chunk $c$ back, it extracts the dream key as the chunk payload $k=\mathit{Data}(c)$.

$D$ retrieves the parity data $C'$ using reference $r$, and then decodes the plaintext as $C=C'\xor k$.
The retrieval process is trivially correct as long as: 
\begin{enumerate}[noitemsep]
\item The parity data $C'$ is retrievable using standard retrieval methods.
\item The dream protocol is followed by all cooperating nodes.
\item The contents of the batch buckets along the dream path all remain unchanged. 
\end{enumerate}

% \subsection{Deletion}
% \label{def:deletion}

We are now turning to access revocation, i.e. deleting content by making it inaccessible.

Uploader $U$ had granted downloader $D$ at address $a$ access to content $C$ through the dream reference $\langle r,b,g\rangle $. $U$ revokes $D$'s access by uploading extra chunks to the buckets batch $b$  designated by the input chunk address ($\mathit{Address}(c_i)\text{ for }0\le i\le n$). 
As a result, downloader $D$ will no longer be able to retrieve content $C$. Because the output of the key update function has changed, the subsequent chunk changes and the dream path is diverted. Since we cannot know which nodes are colluding and malicious, the uploader needs to upload batch bucket changes to all neighbourhoods on the dream path to reliably revoke access.
%
% \subsection{Incentives}

% Discuss the reliability of the protocol in the context of economic incentives.
% \begin{figure}[htbp]
%   \centering
%     \includegraphics[width=0.8\textwidth]{figs/push-payment.pdf}
%   \caption{Incentives for push syncing.}
% \label{fig:push-payment}
% \end{figure}

% How to determine when to forward.
% discuss the cost of a dream

\section{Security}
\label{sec:security}

The reference to a dream chunk \textit{does not leak} any information about the step-count or the neighbourhoods involved. Neither are the participants in the protocol aware of their position within the dream path, nor of any details about the other neighbourhoods that are part of it, except for the immediate next one to which they are push-syncing their chunk.

The construction of the dream reference is safely \emph{deniable} as long as no adversary gets hold of the actual encrypted chunk of $C'$: if we consider our sensitive content $C$, and any uncontroversial chunk content $A$, then when creating $C'$, the owner also encrypts $A'=A\xor k$ and uploads it. When asked about $k$, producing $A'$ makes the denial of other content, including $C'$, more plausible. For a scenario where it cannot be assumed that no adversary can get hold of $C'$, both $C$ and $A$ can be asymmetrically encrypted with a public key of the grantee before being encrypted with $k$, to prevent patterns showing in $A'\xor C'$ that can help breaking the symmetric encryption and thus prevent successful deletion.%
%
\footnote{An adversary possessing both A' and C' could calculate $A'\xor C'$ to find patterns in the result.}

As Swarm is a decentralised network, individual participants cannot (and should not) be trusted to play by the rules. The built-in incentives are what encourage cooperative behaviour and help maintain a Nash equilibrium. To avoid the situation that a single corrupt node in a neighbourhood could keep deleted chunks accessible, the mechanism for deletion, i.e. the strategic update to the relevant bucket, should be applied to every neighbourhood of a dream path. A sufficient condition for content deletion is then that in \emph{at least one neighbourhood} all nodes are honest, i.e. will not store and serve information that they should not.% 
%
\footnote{With the growth of the Swarm network, the number of neighbourhoods increases whereas the membership size per neighbourhoods remains about the same. Thus, the costs of attacking the proposed protocol increases with network size.}
%

Alternatively, a powerful adversary could potentially infiltrate every neighbourhood of Swarm and archive all information that has ever been uploaded (without being able to decipher it). It could also keep logs of what was uploaded in what order, which would allow it to serve `deleted' content. However, there is a huge cost associated with such indiscriminate archiving of all of Swarm's content, in what is the only way to reliably defeat the dream construction. Because the store would also contain all downloads, it is equivalent to the scenario where an adversary makes deletion redundant by keeping all downloaded data, which is a case not commonly interpreted as defeating deletion.

% \subsection{Step Count}
% \label{def:step-count}

We now turn to the discussion of how to calibrate the step count in relation to the security model using the notion of a \emph{network-wide neighbourhood infiltration rate}. For instance, a rate of $\frac{1}{2}$ means that the chance of infiltration, in any neighbourhood independently and uniformly across the network, is 1 in 2, implying that, on average, we can expect half of the neighbourhoods in the network to contain a malicious and colluding node.

For access to be safely revoked, the owner uploads new chunks in each neighbourhood along the dream path. In practice, the malicious node operators would not be able to tell which chunks are strategic updates to facilitate deletion. But we will assume for the sake of the argument that malicious nodes are able and willing to disregard these new chunks and, in an attempt to render the revocation of access impotent, respond to a request the same as before. If every neighbourhood on the path contains at least one malicious node, it is, in principle, possible for the dream protocol to return the unchanged response, and access is granted against the intention of the uploader, i.e. the protocol terminates allowing access even though it was revoked. Given a specific dream path, however, if even a single neighbourhood on the dream path is honest, they will respect the chunks that newly arrived since the construction and therefore, divert the dream path, preventing it from terminating with the grantee.

Thus, for a breach of access to happen, all neighbourhoods must be malicious.\footnote{This is a necessary condition, but not a sufficient one: even a malicious neighbourhood could act honestly, if the route happens to pass through one of its honest nodes. This further, significant improvement of the odds that deletion succeeds is for simplicity ignored in the following. }
In our security model, a neighbourhood is malicious with uniform and independent probability.  For an overall infiltration rate of 1 out of $k$, the chance of all neighbourhoods on a given random dream path being malicious is $k^{-n}$. For a security requirement of a success rate of $\sigma$ "nines", i.e. for an error rate less than $10^{-\sigma}$, we can formulate the requirement as
\begin{equation}
k^{-n}\leq10^{-\sigma}
\end{equation}
Now, expressing $k$ as a power of 10, $k=10^\kappa$, taking the logarithm of both sides and multiplying by $-1$, we get
\begin{equation}
    n \geq \frac{\sigma}{\kappa}
\end{equation}
As an example, with one in every 10 neighbourhoods being malicious ($\kappa=1$), we calibrate $n\geq\sigma$: the dream path must be as long as the number of nines expressing the desired success rate.%
%
\footnote{a 6-hop-long path will guarantee access revocation with a certainty of $99.9999\%$.}

\section{Implementation Constraints}\label{sec:constraints}

Dream can be implemented on the existing Swarm network with only minor additions and modifications to the DISC protocol, its proven data storage mechanism. Because the dream protocol is simple and strategically utilises the DISC primitives, the effort required to implement it on Swarm would be modest.

Beyond its technical capacity, the size, speed, proven availability and maturity of the Swarm network\footnote{See the Swarm dashboard at \url{https://network.ethswarm.org/} and the monitoring tool called Swarm Scan at \url{https://swarmscan.io/}} also make it suitable to serve as the underlying layer for the dream protocol. At a size of almost 10,000 nodes and a global spread across four continents, the network is large and distributed enough to provide the maze of nodes needed to make dream key creation and retrieval secure, and re-creation after intended deletion practically impossible.

The reference used by the grantees to retrieve content is more complex then Swarm's standard reference. Analogous to the way encryption is handled, the length of the reference allows to identify the referenced content as deletable.

\section{Conclusion}\label{sec:conclusion}
Based on the concept of revocable access, this paper presents a technically novel yet practical mechanism to achieve deletable content in the immutable context of peer-to-peer storage. By formalising deletion as the loss of access rather than the physical erasure of data, the dream construct offers a secure, permissionless, and self-sovereign solution. The system is designed to integrate seamlessly with Swarm’s DISC storage model and relies on simple yet robust protocol interactions among peer nodes. Through the introduction of a distributed key and a network-driven computation model, dream enables users to selectively revoke access to content without undermining the foundational principles of decentralisation, immutability, and censorship resistance. The result is a useful balance between data permanence and sovereign control, which offers a much-needed, viable approach to privacy-conscious publishing in the Web3 era.

The ability to delete data from immutable storage marks an essential contribution towards maturity for decentralised storage technology. It adds a feature that centralised, mutable systems so far seemed to have as a permanent advantage over the alternatives. Once established as a standard mode of storage, this can reduce regulatory friction and empower users. It could enable a new class of applications to utilise decentralised storage, which depend on compliance to be able to operate; and it caters to privacy-conscious users who wish stronger control over the data they share.
While DREAM realises a subset of the `right to be forgotten.' The fact that Swarm resists censorship while enabling deletion points to a promising middle ground as a potential direction for regulators of commercial data hubs.

The way that DREAM uses Swarm's network topology to `store' a key–and the fact that a simple upload can change the conditions to effectively revoke access—may serve as an inspiration to rethink other seemingly intractable problems of web3.

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
The authors would like to thank György Barabás, Andrea Robert for help with text and formatting, Daniel Nickless for figure \ref{fig:pushsync}, and members of the Swarm Research Division, especially Daniel A.~Nagy for their ideas leading to this paper.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}

\endinput
